---
title: "Analysis"
author: "Hanan Ali"
date: "2025-06-25"
output: html_document
---

Employer dataset prep
```{r setup, include=FALSE}
######################################
##           Load Library           ##
######################################

#Complimentary "Install Packages" code
#install.packages("tidyverse")
#install.packages("janitor")
#install.packages("gtools")
#install.packages("naniar")

#Load Libraries
library(tidyverse)
library(janitor)
library(gtools)
library(naniar)


#############################################
## Connect, Import, group and primary key. ##
#############################################

#Read data into R environment. 
Employer_Data <- read_csv("Employer_Dentistry_Survey.csv") %>% clean_names()
#Name new object to work with leaving original intact/untouched
Employer_Data_Clean <- Employer_Data

#Add grouping identifier now so if we combine data later, we have our groups delineated. 
Employer_Data_Clean <- Employer_Data_Clean %>%
  mutate(
    group = "Employer", #Categorizes all responses as Employer
    respondent_id = row_number() #Maintains respondent ID
  )


############################
#####  DATA WRANGLING  #####
############################

#Adjust headers for readability and add grouping identifier.
#1.) Remove redundancy in column names, 
#2.) Replace "quid" with "Q", and
#3.) All "_text" with "T".
colnames(Employer_Data_Clean) <- gsub("_text$", "_T", gsub("q", "Q", colnames(Employer_Data_Clean)))
#4.) Format the column headers so that it is sort-able both by question and sub question. So where 
#questions are formatted with header Q10_2, the header will now be Q10_02.  This preserves ordering
#and makes indexing easier. 
pad_question_ids <- function(col_names) {
  col_names %>%
    str_replace_all("Q(\\d{1})(?!\\d)", "Q0\\1") %>%
    str_replace_all("Q(\\d{2})_(\\d{1})(?!\\d)", "Q\\1_0\\2")}

#Apply above formatting changes to data frame column names
colnames(Employer_Data_Clean) <- pad_question_ids(colnames(Employer_Data_Clean))

#Removes Qualtric metadata from the first two rows
Employer_Data_Clean <- Employer_Data_Clean %>% slice(-1, -2)



### QUESTION-BY-QUESTION Cleaning

#------------
#Question #04
#------------
#Renames the text response to index 13 and places it in its order. 
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q04_13 = Q04_12_T)

#------------
#Question #09 does not need correcting. 
#------------

#------------
#Question #16
#------------
#Renames the text response to index 11 and places it in its order. 
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q16_11 = Q16_10_T) 

#Process Q12: Split responses and pivot to wide format
Employer_Data_Clean <- Employer_Data_Clean %>%
  mutate(respondent_id = row_number()) %>%
  separate_rows(Q16, sep = ",") %>%
  mutate(Q16 = str_trim(Q16)) %>%
  #filter(!is.na(Q16)) %>%  # 🔧 prevent Q16_NA column in pivot_wider
  mutate(Q16 = str_pad(Q16, width = 2, pad = "0"),
         value = 1) %>%
  pivot_wider(
    names_from = Q16,
    names_prefix = "Q16_",
    values_from = value,
    values_fill = 0
  ) %>%
  right_join(select(Employer_Data_Clean, respondent_id), by = "respondent_id")  # 🔁 bring back NA responders


if (!"Q16_09" %in% colnames(Employer_Data_Clean)) {
  Employer_Data_Clean$Q16_09 <- 0
}


#------------
#Question #17
#------------
#Renames the text response to index 14 and places it in its order.
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q17_15 = Q17_14_T)

#------------
#Question #20
#------------
# Create binary indicator columns from Q16.  Recall that Q16 was a multiple responsive
# "Select all that Apply" question where participants could choose however many they 
# wanted.  In the Qualtrics output, these responses were comma-separated.  This query
# breaks each reponse out over a column and assigns values of "0" or "1" depending on 
# whether they choose that particular response. The same issue is address in Question #20. 
Q20_binary <- Employer_Data_Clean %>%
  select(respondent_id, Q20) %>%
  separate_rows(Q20, sep = ",") %>%
  mutate(Q20 = str_trim(Q20)) %>%
  #filter(!is.na(Q20)) %>%   # 👈 Prevents Q20_NA from being created
  mutate(Q20 = str_pad(Q20, width = 2, pad = "0"),
         value = 1) %>%
  pivot_wider(
    names_from = Q20,
    names_prefix = "Q20_",
    values_from = value,
    values_fill = 0
  ) %>%
  right_join(select(Employer_Data_Clean, respondent_id), by = "respondent_id")

# Remove the old Q16 columns and join the binary ones on respondent_id
Employer_Data_Clean <- Employer_Data_Clean %>%
  select(-Q20) %>%
  left_join(Q20_binary, by = "respondent_id")

# Renames the text response to index 07 and places it in its order.
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q20_07 = Q20_06_T) %>%
  relocate(Q20_07, .after = Q20_06)

# Add Q16_05 as a column of zeros if it doesn't exist (it does not)
if (!"Q20_05" %in% colnames(Employer_Data_Clean)) {
  Employer_Data_Clean$Q20_05 <- 0
}

#------------
#Question #21
#------------
#Renames the text response to index 07 and places it in its order.
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q21_07 = Q21_06_T) 

#------------
#Question #22 does not need correcting. 
#------------

#------------
#Question #24 does not need correcting. 
#------------

#------------
#Question #25
#------------
# Create binary indicator columns from Q25.  Recall that Q25 was a multiple responsive
# "Select all that Apply" question where participants could choose however many they 
# wanted.  In the Qualtrics output, these responses were comma-separated.  This query
# breaks each reponse out over a column and assigns values of "0" or "1" depending on 
# whether they choose that particular response. The same issue is address in Question #16. 
Q25_binary <- Employer_Data_Clean %>%
  select(respondent_id, Q25) %>%
  separate_rows(Q25, sep = ",") %>%
  mutate(Q25 = str_trim(Q25)) %>%
  #filter(!is.na(Q25)) %>%   # 👈 Prevents Q25_NA
  mutate(Q25 = str_pad(Q25, width = 2, pad = "0"),
         value = 1) %>%
  pivot_wider(
    names_from = Q25,
    names_prefix = "Q25_",
    values_from = value,
    values_fill = 0
  ) %>%
  right_join(select(Employer_Data_Clean, respondent_id), by = "respondent_id")


# Remove the old Q16 columns and join the binary ones in
Employer_Data_Clean <- Employer_Data_Clean %>%
  select(-Q25) %>%
  left_join(Q25_binary, by = "respondent_id")

# Add Q25_13 as a column of zeros if it doesn't exist (it does not)
if (!"Q25_13" %in% colnames(Employer_Data_Clean)) {
  Employer_Data_Clean$Q25_13 <- 0
}

# Rename and relocate the "Other" text entry for Q16
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q25_14 = Q25_13_T) %>%
  relocate(Q25_14, .after = Q25_13)

#------------
#Question #26
#------------
#Renames the text response to index 14 and places it in its order. 
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q26_14 = Q26_13_T) 


##########################
##### FILTER DATASET #####
##########################
#"Q20_NA"
#"Q25_NA"


#Remove irrelevant/problematic columns to the research/stats questions
Employer_Data_Clean <- Employer_Data_Clean %>%
  mutate(progress = as.numeric(progress)) %>%
  filter(Q02 == '1', progress >= 50) %>%
  select(
    ip_address, progress, duration_in_seconds,
    finished, recorded_date, response_id, group, respondent_id,
    starts_with("Q04"),
    starts_with("Q09"),
    starts_with("Q16"),
    starts_with("Q17"),
    starts_with("Q20"),
    starts_with("Q21"),
    starts_with("Q22"),
    starts_with("Q24"),
    starts_with("Q25"),
    starts_with("Q26"),
    -Q16_NA,  # 👈 drop this
    -Q20_NA,  # 👈 drop this
    -Q25_NA
  )

#Identify metadata columns to keep at the front
meta_cols <- c("ip_address", "progress", "duration_in_seconds",
               "finished", "recorded_date", "response_id", "group", "respondent_id")

#Identify and sort Q columns using mixedsort
q_cols <- setdiff(colnames(Employer_Data_Clean), meta_cols)
q_cols_sorted <- mixedsort(q_cols)

#Reorder the data frame
Employer_Data_Clean <- Employer_Data_Clean[, c(meta_cols, q_cols_sorted)]

#Print the entire dataframe
Employer_Data_Clean %>% print(n = 43)
colnames(Employer_Data_Clean)


count(Employer_Data_Clean)
```



Educator Survey Dataset Prep
```{r}
######################################
##           Load Library           ##
######################################

#Complimentary "Install Packages" code
#install.packages("tidyverse")
#install.packages("janitor")
#install.packages("gtools")
#install.packages("naniar")

#Load Libraries
library(tidyverse)
library(janitor)
library(gtools)
library(naniar)


#############################################
## Connect, Import, group and primary key. ##
#############################################

#Read data into R environment. 
Educator_Data <- read_csv("PCVE_Dentistry_Survey.csv") %>% clean_names()
#Name new object to work with leaving original intact/untouched
Educator_Data_Clean <- Educator_Data

#Add identifier affirming educator responses. 
Educator_Data_Clean <- Educator_Data_Clean %>%
  mutate(
    group = "Educator",
    respondent_id = row_number()
  )

############################
#####  DATA WRANGLING  #####
############################

#Adjust headers for readability and add grouping identifier.
  #1.) Remove redundancy in column names, 
  #2.) Replace "quid" with "Q", and
  #3.) All "_text" with "T".
colnames(Educator_Data_Clean) <- gsub("_text$", "_T", gsub("q", "Q", colnames(Educator_Data_Clean)))
  #4.) Format the column headers so that it is sort-able both by question and sub question. So where 
    #questions are formatted with header Q10_2, the header will now be Q10_02.  This preserves ordering
    #and makes indexing easier. 
pad_question_ids <- function(col_names) {
  col_names %>%
    str_replace_all("Q(\\d{1})(?!\\d)", "Q0\\1") %>%
    str_replace_all("Q(\\d{2})_(\\d{1})(?!\\d)", "Q\\1_0\\2")}

#Apply above formatting changes to data frame column names
colnames(Educator_Data_Clean) <- pad_question_ids(colnames(Educator_Data_Clean))

#Removes Qualtric metadata from the first two rows
Educator_Data_Clean <- Educator_Data_Clean %>% slice(-1, -2)


### QUESTION-BY-QUESTION Cleaning

#------------
#Question #04
#------------
#Renames the text response to index 13 and places it in its order. 
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q04_13 = Q04_12_T)

#------------
#Question #07 does not need correcting. 
#------------

#------------
#Question #12
#------------
  #Renames the text response to index 11 and places it in its order. 
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q12_11 = Q12_10_T) 

  #Process Q12: Split responses and pivot to wide format
Educator_Data_Clean <- Educator_Data_Clean %>%
  # Create a temporary respondent ID if not already done
  mutate(respondent_id = row_number()) %>%
  # Separate comma-separated values into long format
  separate_rows(Q12, sep = ",") %>%
  # Trim whitespace if any
  mutate(Q12 = str_trim(Q12)) %>%
  # Pad values with leading zero for consistent column naming (01, 02, ..., 11)
  mutate(Q12 = str_pad(Q12, width = 2, pad = "0"),
         value = 1) %>%
  # Pivot wider to get binary indicator columns
  pivot_wider(
    names_from = Q12,
    names_prefix = "Q12_",
    values_from = value,
    values_fill = 0
  )

#------------
#Question #13
#------------
#Renames the text response to index 14 and places it in its order.
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q13_14 = Q13_13_T)

#------------
#Question #16
#------------
# Create binary indicator columns from Q16.  Recall that Q16 was a multiple responsive
# "Select all that Apply" question where participants could choose however many they 
# wanted.  In the Qualtrics output, these responses were comma-separated.  This query
# breaks each reponse out over a column and assigns values of "0" or "1" depending on 
# whether they choose that particular response. The same issue is address in Question #20. 
Q16_binary <- Educator_Data_Clean %>%
  select(respondent_id, Q16) %>%
  separate_rows(Q16, sep = ",") %>%
  mutate(Q16 = str_trim(Q16)) %>%
  mutate(Q16 = str_pad(Q16, width = 2, pad = "0"),
         value = 1) %>%
  pivot_wider(
    names_from = Q16,
    names_prefix = "Q16_",
    values_from = value,
    values_fill = 0
  )

  # Remove the old Q16 columns and join the binary ones on respondent_id
Educator_Data_Clean <- Educator_Data_Clean %>%
  select(-Q16) %>%
  left_join(Q16_binary, by = "respondent_id")

  # Renames the text response to index 07 and places it in its order.
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q16_07 = Q16_06_T) %>%
  relocate(Q16_07, .after = Q16_06)

  # Add Q16_05 as a column of zeros if it doesn't exist (it does not)
if (!"Q16_05" %in% colnames(Educator_Data_Clean)) {
  Educator_Data_Clean$Q16_05 <- 0
}

#------------
#Question #17
#------------
#Renames the text response to index 07 and places it in its order.
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q17_07 = Q17_06_T) 

#------------
#Question #18 does not need correcting. 
#------------

#------------
#Question #19 does not need correcting. 
#------------

#------------
#Question #20
#------------
# Create binary indicator columns from Q16.  Recall that Q16 was a multiple responsive
# "Select all that Apply" question where participants could choose however many they 
# wanted.  In the Qualtrics output, these responses were comma-separated.  This query
# breaks each reponse out over a column and assigns values of "0" or "1" depending on 
# whether they choose that particular response. The same issue is address in Question #16. 
Q20_binary <- Educator_Data_Clean %>%
  select(respondent_id, Q20) %>%
  separate_rows(Q20, sep = ",") %>%
  mutate(Q20 = str_trim(Q20)) %>%
  mutate(Q20 = str_pad(Q20, width = 2, pad = "0"),
         value = 1) %>%
  pivot_wider(
    names_from = Q20,
    names_prefix = "Q20_",
    values_from = value,
    values_fill = 0
  )

# Remove the old Q16 columns and join the binary ones in
Educator_Data_Clean <- Educator_Data_Clean %>%
  select(-Q20) %>%
  left_join(Q20_binary, by = "respondent_id")

# Rename and relocate the "Other" text entry for Q16
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q20_14 = Q20_13_T) %>%
  relocate(Q20_14, .after = Q20_13)

#------------
#Question #21
#------------
#Renames the text response to index 14 and places it in its order. 
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q21_14 = Q21_13_T) 


##########################
##### FILTER DATASET #####
##########################

#Remove irrelevant/problematic columns to the research/stats questions
Educator_Data_Clean <- Educator_Data_Clean %>%
  mutate(progress = as.numeric(progress)) %>%
  filter(Q44 != "2", progress >= 50) %>%
  select(
    ip_address, progress, duration_in_seconds,
    finished, recorded_date, response_id, group, respondent_id,
    starts_with("Q04"),
    starts_with("Q07"),
    starts_with("Q12"),
    starts_with("Q13"),
    starts_with("Q16"),
    starts_with("Q17"),
    starts_with("Q18"),
    starts_with("Q19"),
    starts_with("Q20"),
    starts_with("Q21")
  )



#Identify metadata columns to keep at the front
meta_cols <- c("ip_address", "progress", "duration_in_seconds",
               "finished", "recorded_date", "response_id", "group", "respondent_id")

#Identify and sort Q columns using mixedsort
q_cols <- setdiff(colnames(Educator_Data_Clean), meta_cols)
q_cols_sorted <- mixedsort(q_cols)

#Reorder the data frame
Educator_Data_Clean <- Educator_Data_Clean[, c(meta_cols, q_cols_sorted)]
Educator_Data_Clean
colnames(Educator_Data_Clean)

#Print the entire dataframe
Educator_Data_Clean %>% print(n = 43)

count(Educator_Data_Clean)
```




Are there significant differences between educators and practice owners in their belief that new graduates are competent in key dental skills on their first day of practice?

First I'll combine the datasets and keep it to question 4 only
```{r}
library(dplyr)
   
# Add group identifiers
educator_cleaned <- Educator_Data_Clean %>%
  mutate(group = "Educator")

employer_cleaned <- Employer_Data_Clean %>%
  mutate(group = "Employer")

# Define Q4 variable names for both datasets
q4_vars <- paste0("Q04_", sprintf("%02d", 1:12))  # creates Q04_01 to Q04_12

# Subset and label the educator data
educator_q4 <- educator_cleaned %>%
  select(all_of(q4_vars)) %>%
  mutate(group = "Educator")

# Subset and label the employer data
employer_q4 <- employer_cleaned %>%
  select(all_of(q4_vars)) %>%
  mutate(group = "Employer")

# Combine datasets
combined_data <- bind_rows(educator_q4, employer_q4)

```



Run Mann-Whitney U Test
```{r}
library(tidyr)
library(dplyr)
library(purrr)

# Pivot data to long format
q4_long <- combined_data %>%
  pivot_longer(
    cols = starts_with("Q04_"),
    names_to = "Skill",
    values_to = "Rating_num"
  ) %>%
  filter(!is.na(Rating_num))

#Must coerce character to numeric explicitly because removing na values is not enough so redo and see if it works
q4_long <- q4_long %>%
  mutate(Rating_num = as.numeric(Rating_num))

str(q4_long)

# Run Mann-Whitney U test for each skill
mw_results <- q4_long %>%
  filter(!is.na(Rating_num)) %>%  # Exclude rows with NA ratings because there is NA values hindering it from being able to run 
  group_by(Skill) %>%
  summarise(
    p_value = wilcox.test(Rating_num ~ group)$p.value,
    median_educator = median(Rating_num[group == "Educator"]),
    median_employer = median(Rating_num[group == "Employer"]),
    n_educator = sum(group == "Educator"),
    n_employer = sum(group == "Employer"),
    .groups = "drop"
  ) %>%
  arrange(p_value)

print(mw_results)
   

```
We used the Mann-Whitney U test, a non-parametric method appropriate for comparing ordinal ratings between two independent groups. Ratings ranged from 1 (Strongly Disagree) to 4 (Strongly Agree) regarding new graduate competence for each skill.

Results:

For most dental skills, the differences in ratings between educators and practice owners were not statistically significant (p-values well above 0.05), indicating that both groups generally agree on the competency level of new graduates. The skill “interpreting dental radiographs” (dental X-rays) showed the closest difference with a p-value of 0.064. Although this is just above the conventional 0.05 threshold, it suggests a trend where practice owners were slightly more confident in new graduates’ abilities compared to educators. Median ratings for this skill were equal (3 out of 4), reinforcing that the difference is subtle.

Interpretation: There is no strong evidence that educators and practice owners disagree on the preparedness of new graduates for dental tasks. Both groups mostly agree on which skills new graduates can perform competently from day one. The only exception might be interpreting dental X-rays, where practice owners seem a little more optimistic about new grads' skills — but this difference isn’t strong enough to be sure it’s real. The near-significant difference in confidence regarding interpreting dental X-rays may warrant further investigation but should be interpreted cautiously.


 






Is there a difference between educators and practice owners in their reports (educators’ actual teaching vs. owners’ perceptions) of which dental skills were taught in the pre-clinical DVM curriculum for recent graduates?

Employer Q16: “What format of instruction in dentistry does your DVM program provide during the clinical year?” (select all that apply)

Educator Q12: “Which of the following skills are taught as part of the pre-clinical dentistry courses in your DVM program?” (select all that apply)
```{r}
library(dplyr)
library(tidyr)

# 1. Educator Q12 data (Q12_01 to Q12_07)
educator_q12 <- Educator_Data_Clean %>%
  select(Q12_01:Q12_07) %>%
  mutate(group = "Educator")

colnames(educator_q12)[1:7] <- paste0("Q12_", sprintf("%02d", 1:7))

# 2. Employer Q16 data (only Q16_01 to Q16_10)
employer_q16 <- Employer_Data_Clean %>%
  select(Q16_01:Q16_10) %>%  # select only first 10 Q16 columns
  mutate(group = "Employer")

colnames(employer_q16)[1:10] <- paste0("Q16_", sprintf("%02d", 1:10))

# 3. Combine datasets
combined_raw <- bind_rows(educator_q12, employer_q16)

# 4. Convert Q columns to character to avoid pivot issues
combined_raw_clean <- combined_raw %>%
  mutate(across(starts_with("Q"), as.character))

# 5. Pivot longer
combined_long <- combined_raw_clean %>%
  pivot_longer(
    cols = starts_with("Q"),
    names_to = "Skill_Format",
    values_to = "Taught_raw"
  ) %>%
  mutate(
    Taught = na_if(Taught_raw, "N/A"),
    Taught = na_if(Taught, ""),
    Taught = as.numeric(Taught)
  ) %>%
  select(Skill_Format, group, Taught)

# 6. Confirm columns included
combined_long %>%
  filter(grepl("^Q16_", Skill_Format)) %>%
  distinct(Skill_Format)




combined_long %>%
  filter(grepl("^Q16_", Skill_Format)) %>%
  group_by(Skill_Format) %>%
  summarise(
    total_responses = n(),
    missing = sum(is.na(Taught)),
    non_missing = total_responses - missing,
    prop_non_missing = non_missing / total_responses
  )


```

During the analysis of the combined dataset comparing educators’ reports and practice owners’ perceptions of which dental skills are taught in the pre-clinical DVM curriculum, we encountered a data quality issue with the survey item labeled Q12_11. Specifically, all responses for this skill were recorded as missing values (NAs) across both groups—educators and employers. Because there were no actual observed responses for this item, the contingency table used for statistical testing contained no valid data points. Removing Q12_11 allows us to focus on skills where sufficient data exist to reliably assess differences in perceptions about curriculum content between educators and practice owners.

Another item worth pointing out is that the Q16 fields (Q16_01 to Q16_10), out of 43 total responses, approximately 70% are missing, with only about 30% of responses containing usable data. This high level of missingness may limit the robustness of any analysis on these items and should be carefully considered when interpreting results. Further investigation into the reasons for missing data is recommended before proceeding.







Another key item to keep in mind as to why the code above won't work is that the key issue is that the Educator survey uses Q12 questions, while the Employer survey uses Q16 questions for related skills. As a result, the Educator group only has data for Q12 items, and the Employer group only has data for Q16 items. Because the question labels differ, no single skill label contains responses from both groups, preventing valid 2×2 contingency tables needed for Fisher’s exact tests. To compare the two groups on the same skills, we rename Employer Q16 columns to match the Q12 labels. This alignment allows us to combine data by common skill labels, ensuring each skill includes responses from both groups and enabling meaningful statistical comparisons.



```{r}
library(dplyr)
library(tidyr)

# 1. Rename employer Q16 columns to Q12 labels for matching skills
employer_q16 <- Employer_Data_Clean %>%
  select(Q16_01:Q16_07) %>%
  rename_with(~ paste0("Q12_", sprintf("%02d", 1:7)), .cols = everything()) %>%
  mutate(group = "Employer")

# 2. Select educator Q12 columns and label group
educator_q12 <- Educator_Data_Clean %>%
  select(Q12_01:Q12_07) %>%
  mutate(group = "Educator")

# 3. Combine the two datasets
combined_q12_q16 <- bind_rows(educator_q12, employer_q16)

# 4. Convert Q columns to character (to avoid pivot issues)
combined_raw_clean <- combined_q12_q16 %>%
  mutate(across(starts_with("Q12_"), as.character))

# 5. Pivot to long format
combined_long <- combined_raw_clean %>%
  pivot_longer(
    cols = starts_with("Q12_"),
    names_to = "Skill_Format",
    values_to = "Taught_raw"
  ) %>%
  mutate(
    Taught = na_if(Taught_raw, "N/A"),
    Taught = na_if(Taught, ""),
    Taught = as.numeric(Taught)
  ) %>%
  filter(!is.na(Taught), group %in% c("Educator", "Employer"))

# 6. Run Fisher's exact test on valid 2x2 tables
test_results <- combined_long %>%
  group_by(Skill_Format) %>%
  summarise(
    tab = list(table(group, Taught)),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    dims = list(dim(tab)),
    is_valid = !is.null(dims[[1]]) && all(dims[[1]] >= 2)
  ) %>%
  filter(is_valid) %>%
  mutate(
    p_value = tryCatch(fisher.test(tab)$p.value, error = function(e) NA_real_),
    n_educator = if ("Educator" %in% rownames(tab)) sum(tab["Educator", ]) else NA_integer_,
    n_employer = if ("Employer" %in% rownames(tab)) sum(tab["Employer", ]) else NA_integer_,
    prop_educator = if ("Educator" %in% rownames(tab) && "1" %in% colnames(tab)) {
      tab["Educator", "1"] / sum(tab["Educator", ])
    } else NA_real_,
    prop_employer = if ("Employer" %in% rownames(tab) && "1" %in% colnames(tab)) {
      tab["Employer", "1"] / sum(tab["Employer", ])
    } else NA_real_
  ) %>%
  ungroup() %>%
  select(-tab, -is_valid, -dims) %>%
  arrange(p_value)

print(test_results)

```

The Fisher’s exact test results show no statistically significant differences between Educators and Employers across all seven skills assessed (all p-values > 0.32). Both groups reported similar proportions of positive responses, with Educator positive response rates ranging from about 53% to 77%, and Employer rates ranging from about 54% to 92%. These findings suggest that Educators and Employers have broadly comparable perceptions or reports regarding training or expectation of these veterinary dentistry skills. The lack of significant difference implies alignment between the two groups on these specific competencies.






To assess whether educators and employers differ in their beliefs about which dental skills should be included in the pre-clinical DVM curriculum, we analyzed responses to parallel Likert-scale questions: Q13 (educators) and Q17 (employers). Both groups rated the importance of teaching 12 specific dentistry skills using a 4-point Likert scale ranging from 1 (Strongly Disagree) to 4 (Strongly Agree). Because the data were ordinal and responses came from two independent groups, the Mann-Whitney U test was used to compare educator and employer ratings for each skill individually.

The results showed statistically significant differences in opinion for two of the twelve skills. Employers were significantly more likely than educators to agree that feline open extractions involving multiple roots or canines (Skill_11, p = 0.003) and fluoride treatment (Skill_12, p = 0.013) should be included in the pre-clinical curriculum. For all other skills, there were no statistically significant differences between groups (all p > 0.10), and the median rating for most skills was 4 (“Strongly Agree”) in both groups. These findings suggest general alignment between educators and employers on most skills, but highlight a few areas—particularly more advanced feline extractions and fluoride use—where employer expectations may exceed what educators currently prioritize for pre-clinical training.




```{r}
# Load required packages
library(dplyr)
library(tidyr)
library(gt)

# 1. Select relevant Q13 (educators) and Q17 (employers) columns
q13_vars <- paste0("Q13_", sprintf("%02d", 1:12))  # Educators
q17_vars <- paste0("Q17_", sprintf("%02d", 1:12))  # Employers

educator_q13 <- Educator_Data_Clean %>%
  select(all_of(q13_vars)) %>%
  mutate(group = "Educator")

employer_q17 <- Employer_Data_Clean %>%
  select(all_of(q17_vars)) %>%
  mutate(group = "Employer")

# 2. Combine and standardize column names
colnames(educator_q13)[1:12] <- paste0("Skill_", sprintf("%02d", 1:12))
colnames(employer_q17)[1:12] <- paste0("Skill_", sprintf("%02d", 1:12))

combined_data <- bind_rows(educator_q13, employer_q17)

# 3. Reshape to long format
q3_long <- combined_data %>%
  pivot_longer(
    cols = starts_with("Skill_"),
    names_to = "Skill",
    values_to = "Rating_num"
  ) %>%
  mutate(Rating_num = as.numeric(Rating_num)) %>%
  filter(!is.na(Rating_num))

# 4. Run Mann-Whitney U tests
q3_results <- q3_long %>%
  group_by(Skill) %>%
  summarise(
    p_value = wilcox.test(Rating_num ~ group)$p.value,
    median_educator = median(Rating_num[group == "Educator"], na.rm = TRUE),
    median_employer = median(Rating_num[group == "Employer"], na.rm = TRUE),
    n_educator = sum(group == "Educator"),
    n_employer = sum(group == "Employer"),
    .groups = "drop"
  ) %>%
  arrange(p_value)

# 5. Clean column names
q3_results_clean <- q3_results %>%
  rename(
    `p-value` = p_value,
    `Median (Educator)` = median_educator,
    `Median (Employer)` = median_employer,
    `n (Educator)` = n_educator,
    `n (Employer)` = n_employer
  )

# 6. Display results as table
q3_results_clean %>%
  gt() %>%
  tab_header(
    title = "Mann-Whitney U Test: Pre-Clinical Skill Training Beliefs"
  ) %>%
  fmt_number(columns = c(`p-value`), decimals = 3)

```








To assess whether educators and practice owners differ in their expectations for the number of dental procedures students should complete during clinical training, we compared responses from Question 19 of the educator survey and Question 24 of the employer survey. Respondents reported the number of procedures they believe should be completed across four clinical settings: primary care, dentistry rotations, shelter medicine, and lab animal medicine. Since Shapiro-Wilk tests indicated that the data were not normally distributed in any group-setting combination (all p-values < 0.05), we used the Mann-Whitney U test to compare groups for each setting.

The results showed no statistically significant differences in expectations between educators and employers across all four settings (all p-values > 0.09). However, there were modest differences in median values. For example, in dentistry rotations, employers reported a higher median (6.5 procedures) than educators (4 procedures), though the p-value (0.098) did not reach statistical significance. In contrast, both groups reported low or zero median expectations for procedures in lab animal and shelter medicine settings. These findings suggest general alignment between groups, though some settings—particularly dentistry rotations—may reflect subtle differences in expectations that warrant further exploration.


```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(broom)

# 1. Define variables
educator_vars <- paste0("Q19_0", 1:4)
employer_vars <- paste0("Q24_0", 1:4)
setting_labels <- c(
  "Primary_Care", "Dentistry_Rotation", "Shelter_Medicine", "Lab_Animal"
)

# 2. Select and reshape data
educator_proc <- Educator_Data_Clean %>%
  select(all_of(educator_vars)) %>%
  mutate(group = "Educator")

employer_proc <- Employer_Data_Clean %>%
  select(all_of(employer_vars)) %>%
  mutate(group = "Employer")

# 3. Rename columns to match across datasets
colnames(educator_proc)[1:4] <- setting_labels
colnames(employer_proc)[1:4] <- setting_labels

# 4. Combine into long format
combined_proc <- bind_rows(educator_proc, employer_proc) %>%
  pivot_longer(
    cols = all_of(setting_labels),
    names_to = "Setting",
    values_to = "Procedure_Count"
  ) %>%
  filter(!is.na(Procedure_Count)) %>%
  mutate(Procedure_Count = as.numeric(Procedure_Count))

# 5. Check normality (optional step)
combined_proc %>%
  group_by(Setting, group) %>%
  summarise(
    n = n(),
    shapiro_p = shapiro.test(Procedure_Count)$p.value,
    .groups = "drop"
  )

# 6. Run t-tests per setting
ttest_results <- combined_proc %>%
  group_by(Setting) %>%
  summarise(
    ttest = list(t.test(Procedure_Count ~ group)),
    .groups = "drop"
  ) %>%
  mutate(
    tidy_res = map(ttest, broom::tidy)
  ) %>%
  unnest(tidy_res) %>%
  select(Setting, estimate1, estimate2, statistic, p.value, conf.low, conf.high)

# 7. Since normality is violated, use Mann-Whitney
mw_results <- combined_proc %>%
  group_by(Setting) %>%
  summarise(
    p_value = wilcox.test(Procedure_Count ~ group)$p.value,
    median_educator = median(Procedure_Count[group == "Educator"]),
    median_employer = median(Procedure_Count[group == "Employer"]),
    .groups = "drop"
  )

library(gt)

mw_results %>%
  mutate(
    p_value = round(p_value, 3),
    median_educator = round(median_educator, 1),
    median_employer = round(median_employer, 1)
  ) %>%
  rename(
    `Setting` = Setting,
    `Median (Educator)` = median_educator,
    `Median (Employer)` = median_employer,
    `p-value` = p_value
  ) %>%
  gt() %>%
  tab_header(
    title = md("**Mann-Whitney U Test: Expected Number of Dental Procedures**"),
    subtitle = "Comparison of Educator and Employer Expectations by Clinical Setting"
  ) %>%
  cols_align(align = "center") %>%
  fmt_number(columns = `p-value`, decimals = 3)
```




