---
title: "Bridging the Gap:"
subtitle: "Comparing Employer and Educator Expectations in Small Animal Dentistry"
author: Brock Akerman, Hanan Ali, Taylor Cesarski
date: "July, 28, 2025"
bibliography: references.bib
# csl: apa.csl  # Optional: uncomment and provide a CSL file for citation styling (e.g., APA, MLA)
format:
  pdf:
    pdf-engine: xelatex
    mainfont: "Times New Roman"
    toc: true
    toc-depth: 2
    number-sections: true
    number-figures: true
    fig-caption: true
    number-depth: 2
    fig-align: center
    fig-cap-location: bottom
    fig-pos: 'H'
    geometry: margin=1in
    fontsize: 11pt
    keep-tex: true
    linkcolor: blue

editor_options: 
  chunk_output_type: inline

header-includes:
  - \numberwithin{figure}{section}
  - \numberwithin{table}{section}
  - \usepackage{etoolbox}
  - \makeatletter
  - \patchcmd{\maketitle}{\thispagestyle{plain}}{\thispagestyle{plain}\newpage}{}{}
  - \makeatother
---


```{r package_load, warning=FALSE, include=FALSE}
library(ggplot2)
library(tidyverse)
library(janitor)
library(gtools)
library(naniar)
library(patchwork)
library(maps)
library(knitr)
library(dplyr)
library(tidyr)
library(purrr)
library(effectsize) # For calculating ranked-biserial correlation
library(stringr)    
library(kableExtra) # For advanced kable features like column_spec and styling
library(magrittr)
library(forcats) # For fct_rev and fct_relevel
library(scales)  # For scales::label_percent()
```

```{r Employer Data load, echo=FALSE, message=FALSE, warning=FALSE, results="hide",include=FALSE, error=FALSE}
#############################################
## Connect, Import, group and primary key. ##
#############################################

#Read data into R environment. 
Employer_Data <- read_csv("Employer_Dentistry_Survey.csv") %>% clean_names()
#Name new object to work with leaving original intact/untouched
Employer_Data_Clean <- Employer_Data

#Add grouping identifier now so if we combine data later, we have our groups delineated. 
Employer_Data_Clean <- Employer_Data_Clean %>%
  mutate(
    group = "Employer", #Categorizes all responses as Employer
    respondent_id = row_number() #Maintains respondent ID
  )


############################
#####  DATA WRANGLING  #####
############################

#Adjust headers for readability and add grouping identifier.
#1.) Remove redundancy in column names, 
#2.) Replace "quid" with "Q", and
#3.) All "_text" with "T".
colnames(Employer_Data_Clean) <- gsub("_text$", "_T", gsub("q", "Q", colnames(Employer_Data_Clean)))
#4.) Format the column headers so that it is sort-able both by question and sub question. So where 
#questions are formatted with header Q10_2, the header will now be Q10_02.  This preserves ordering
#and makes indexing easier. 
emp_pad_question_ids <- function(col_names) {
  col_names %>%
    str_replace_all("Q(\\d{1})(?!\\d)", "Q0\\1") %>%
    str_replace_all("Q(\\d{2})_(\\d{1})(?!\\d)", "Q\\1_0\\2")}

#Apply above formatting changes to data frame column names
colnames(Employer_Data_Clean) <- emp_pad_question_ids(colnames(Employer_Data_Clean))

#Removes Qualtric metadata from the first two rows
Employer_Data_Clean <- Employer_Data_Clean %>% slice(-1, -2)



### QUESTION-BY-QUESTION Cleaning

#------------
#Question #04
#------------
#Renames the text response to index 13 and places it in its order. 
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q04_13 = Q04_12_T)

#------------
#Question #09 does not need correcting. 
#------------

#------------
#Question #16
#------------
#Renames the text response to index 11 and places it in its order. 
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q16_11 = Q16_10_T) 

#Process Q12: Split responses and pivot to wide format
Employer_Data_Clean <- Employer_Data_Clean %>%
  mutate(respondent_id = row_number()) %>%
  separate_rows(Q16, sep = ",") %>%
  mutate(Q16 = str_trim(Q16)) %>%
  #filter(!is.na(Q16)) %>%  # 🔧 prevent Q16_NA column in pivot_wider
  mutate(Q16 = str_pad(Q16, width = 2, pad = "0"),
         value = 1) %>%
  pivot_wider(
    names_from = Q16,
    names_prefix = "Q16_",
    values_from = value,
    values_fill = 0
  ) %>%
  right_join(select(Employer_Data_Clean, respondent_id), by = "respondent_id")  # 🔁 bring back NA responders


if (!"Q16_09" %in% colnames(Employer_Data_Clean)) {
  Employer_Data_Clean$Q16_09 <- 0
}


#------------
#Question #17
#------------
#Renames the text response to index 14 and places it in its order.
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q17_15 = Q17_14_T)

#------------
#Question #20
#------------
# Create binary indicator columns from Q16.  Recall that Q16 was a multiple responsive
# "Select all that Apply" question where participants could choose however many they 
# wanted.  In the Qualtrics output, these responses were comma-separated.  This query
# breaks each reponse out over a column and assigns values of "0" or "1" depending on 
# whether they choose that particular response. The same issue is address in Question #20. 
Q20_binary <- Employer_Data_Clean %>%
  select(respondent_id, Q20) %>%
  separate_rows(Q20, sep = ",") %>%
  mutate(Q20 = str_trim(Q20)) %>%
  #filter(!is.na(Q20)) %>%   # 👈 Prevents Q20_NA from being created
  mutate(Q20 = str_pad(Q20, width = 2, pad = "0"),
         value = 1) %>%
  pivot_wider(
    names_from = Q20,
    names_prefix = "Q20_",
    values_from = value,
    values_fill = 0
  ) %>%
  right_join(select(Employer_Data_Clean, respondent_id), by = "respondent_id")

# Remove the old Q16 columns and join the binary ones on respondent_id
Employer_Data_Clean <- Employer_Data_Clean %>%
  select(-Q20) %>%
  left_join(Q20_binary, by = "respondent_id")

# Renames the text response to index 07 and places it in its order.
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q20_07 = Q20_06_T) %>%
  relocate(Q20_07, .after = Q20_06)

# Add Q16_05 as a column of zeros if it doesn't exist (it does not)
if (!"Q20_05" %in% colnames(Employer_Data_Clean)) {
  Employer_Data_Clean$Q20_05 <- 0
}

#------------
#Question #21
#------------
#Renames the text response to index 07 and places it in its order.
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q21_07 = Q21_06_T) 

#------------
#Question #22 does not need correcting. 
#------------

#------------
#Question #24 does not need correcting. 
#------------

#------------
#Question #25
#------------
# Create binary indicator columns from Q25.  Recall that Q25 was a multiple responsive
# "Select all that Apply" question where participants could choose however many they 
# wanted.  In the Qualtrics output, these responses were comma-separated.  This query
# breaks each reponse out over a column and assigns values of "0" or "1" depending on 
# whether they choose that particular response. The same issue is address in Question #16. 
Q25_binary <- Employer_Data_Clean %>%
  select(respondent_id, Q25) %>%
  separate_rows(Q25, sep = ",") %>%
  mutate(Q25 = str_trim(Q25)) %>%
  #filter(!is.na(Q25)) %>%   # 👈 Prevents Q25_NA
  mutate(Q25 = str_pad(Q25, width = 2, pad = "0"),
         value = 1) %>%
  pivot_wider(
    names_from = Q25,
    names_prefix = "Q25_",
    values_from = value,
    values_fill = 0
  ) %>%
  right_join(select(Employer_Data_Clean, respondent_id), by = "respondent_id")


# Remove the old Q16 columns and join the binary ones in
Employer_Data_Clean <- Employer_Data_Clean %>%
  select(-Q25) %>%
  left_join(Q25_binary, by = "respondent_id")

# Add Q25_13 as a column of zeros if it doesn't exist (it does not)
if (!"Q25_13" %in% colnames(Employer_Data_Clean)) {
  Employer_Data_Clean$Q25_13 <- 0
}

# Rename and relocate the "Other" text entry for Q16
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q25_14 = Q25_13_T) %>%
  relocate(Q25_14, .after = Q25_13)

#------------
#Question #26
#------------
#Renames the text response to index 14 and places it in its order. 
Employer_Data_Clean <- Employer_Data_Clean %>%
  rename(Q26_14 = Q26_13_T) 


##########################
##### FILTER DATASET #####
##########################
#"Q20_NA"
#"Q25_NA"


#Remove irrelevant/problematic columns to the research/stats questions
Employer_Data_Clean <- Employer_Data_Clean %>%
  mutate(progress = as.numeric(progress)) %>%
  filter(Q02 == '1', progress >= 50) %>%
  select(
    ip_address, progress, duration_in_seconds,
    finished, recorded_date, response_id, group, respondent_id,
    starts_with("Q04"),
    starts_with("Q09"),
    starts_with("Q16"),
    starts_with("Q17"),
    starts_with("Q20"),
    starts_with("Q21"),
    starts_with("Q22"),
    starts_with("Q24"),
    starts_with("Q25"),
    starts_with("Q26"),
    -Q16_NA,  # 👈 drop this
    -Q20_NA,  # 👈 drop this
    -Q25_NA
  )

#Identify metadata columns to keep at the front
emp_meta_cols <- c("ip_address", "progress", "duration_in_seconds",
               "finished", "recorded_date", "response_id", "group", "respondent_id")

#Identify and sort Q columns using mixedsort
emp_q_cols <- setdiff(colnames(Employer_Data_Clean), emp_meta_cols)
emp_q_cols_sorted <- mixedsort(emp_q_cols)

#Reorder the data frame
Employer_Data_Clean <- Employer_Data_Clean[, c(emp_meta_cols, emp_q_cols_sorted)]
```

```{r Educator Data load, echo=FALSE, message=FALSE, warning=FALSE, results="hide",include=FALSE, error=FALSE}
#############################################
## Connect, Import, group and primary key. ##
#############################################

#Read data into R environment. 
Educator_Data <- read_csv("PCVE_Dentistry_Survey.csv") %>% clean_names()
#Name new object to work with leaving original intact/untouched
Educator_Data_Clean <- Educator_Data

#Add identifier affirming educator responses. 
Educator_Data_Clean <- Educator_Data_Clean %>%
  mutate(
    group = "Educator",
    respondent_id = row_number()
  )

############################
#####  DATA WRANGLING  #####
############################

#Adjust headers for readability and add grouping identifier.
  #1.) Remove redundancy in column names, 
  #2.) Replace "quid" with "Q", and
  #3.) All "_text" with "T".
colnames(Educator_Data_Clean) <- gsub("_text$", "_T", gsub("q", "Q", colnames(Educator_Data_Clean)))
  #4.) Format the column headers so that it is sort-able both by question and sub question. So where 
    #questions are formatted with header Q10_2, the header will now be Q10_02.  This preserves ordering
    #and makes indexing easier. 
edu_pad_question_ids <- function(col_names) {
  col_names %>%
    str_replace_all("Q(\\d{1})(?!\\d)", "Q0\\1") %>%
    str_replace_all("Q(\\d{2})_(\\d{1})(?!\\d)", "Q\\1_0\\2")}

#Apply above formatting changes to data frame column names
colnames(Educator_Data_Clean) <- edu_pad_question_ids(colnames(Educator_Data_Clean))

#Removes Qualtric metadata from the first two rows
Educator_Data_Clean <- Educator_Data_Clean %>% slice(-1, -2)


### QUESTION-BY-QUESTION Cleaning

#------------
#Question #04
#------------
#Renames the text response to index 13 and places it in its order. 
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q04_13 = Q04_12_T)

#------------
#Question #07 does not need correcting. 
#------------

#------------
#Question #12
#------------
  #Renames the text response to index 11 and places it in its order. 
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q12_11 = Q12_10_T) 

  #Process Q12: Split responses and pivot to wide format
Educator_Data_Clean <- Educator_Data_Clean %>%
  # Create a temporary respondent ID if not already done
  mutate(respondent_id = row_number()) %>%
  # Separate comma-separated values into long format
  separate_rows(Q12, sep = ",") %>%
  # Trim whitespace if any
  mutate(Q12 = str_trim(Q12)) %>%
  # Pad values with leading zero for consistent column naming (01, 02, ..., 11)
  mutate(Q12 = str_pad(Q12, width = 2, pad = "0"),
         value = 1) %>%
  # Pivot wider to get binary indicator columns
  pivot_wider(
    names_from = Q12,
    names_prefix = "Q12_",
    values_from = value,
    values_fill = 0
  )

#------------
#Question #13
#------------
#Renames the text response to index 14 and places it in its order.
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q13_14 = Q13_13_T)

#------------
#Question #16
#------------
# Create binary indicator columns from Q16.  Recall that Q16 was a multiple responsive
# "Select all that Apply" question where participants could choose however many they 
# wanted.  In the Qualtrics output, these responses were comma-separated.  This query
# breaks each reponse out over a column and assigns values of "0" or "1" depending on 
# whether they choose that particular response. The same issue is address in Question #20. 
Q16_binary <- Educator_Data_Clean %>%
  select(respondent_id, Q16) %>%
  separate_rows(Q16, sep = ",") %>%
  mutate(Q16 = str_trim(Q16)) %>%
  mutate(Q16 = str_pad(Q16, width = 2, pad = "0"),
         value = 1) %>%
  pivot_wider(
    names_from = Q16,
    names_prefix = "Q16_",
    values_from = value,
    values_fill = 0
  )

  # Remove the old Q16 columns and join the binary ones on respondent_id
Educator_Data_Clean <- Educator_Data_Clean %>%
  select(-Q16) %>%
  left_join(Q16_binary, by = "respondent_id")

  # Renames the text response to index 07 and places it in its order.
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q16_07 = Q16_06_T) %>%
  relocate(Q16_07, .after = Q16_06)

  # Add Q16_05 as a column of zeros if it doesn't exist (it does not)
if (!"Q16_05" %in% colnames(Educator_Data_Clean)) {
  Educator_Data_Clean$Q16_05 <- 0
}

#------------
#Question #17
#------------
#Renames the text response to index 07 and places it in its order.
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q17_07 = Q17_06_T) 

#------------
#Question #18 does not need correcting. 
#------------

#------------
#Question #19 does not need correcting. 
#------------

#------------
#Question #20
#------------
# Create binary indicator columns from Q16.  Recall that Q16 was a multiple responsive
# "Select all that Apply" question where participants could choose however many they 
# wanted.  In the Qualtrics output, these responses were comma-separated.  This query
# breaks each reponse out over a column and assigns values of "0" or "1" depending on 
# whether they choose that particular response. The same issue is address in Question #16. 
Q20_binary <- Educator_Data_Clean %>%
  select(respondent_id, Q20) %>%
  separate_rows(Q20, sep = ",") %>%
  mutate(Q20 = str_trim(Q20)) %>%
  mutate(Q20 = str_pad(Q20, width = 2, pad = "0"),
         value = 1) %>%
  pivot_wider(
    names_from = Q20,
    names_prefix = "Q20_",
    values_from = value,
    values_fill = 0
  )

# Remove the old Q16 columns and join the binary ones in
Educator_Data_Clean <- Educator_Data_Clean %>%
  select(-Q20) %>%
  left_join(Q20_binary, by = "respondent_id")

# Rename and relocate the "Other" text entry for Q16
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q20_14 = Q20_13_T) %>%
  relocate(Q20_14, .after = Q20_13)

#------------
#Question #21
#------------
#Renames the text response to index 14 and places it in its order. 
Educator_Data_Clean <- Educator_Data_Clean %>%
  rename(Q21_14 = Q21_13_T) 


##########################
##### FILTER DATASET #####
##########################

#Remove irrelevant/problematic columns to the research/stats questions
Educator_Data_Clean <- Educator_Data_Clean %>%
  mutate(progress = as.numeric(progress)) %>%
  filter(Q44 != "2", progress >= 50, !(response_id %in% c('R_1By20if00mw1SfP', 'R_7kFWGHJpi0UkB7X', 'R_5LXUELTJZBjMt84','R_4CdkBhi7XaGfwNb','R_8EYfypBzjIWzwn7')) ) %>%
  select(
    ip_address, progress, duration_in_seconds,
    finished, recorded_date, response_id, group, respondent_id,
    starts_with("Q04"),
    starts_with("Q07"),
    starts_with("Q12"),
    starts_with("Q13"),
    starts_with("Q16"),
    starts_with("Q17"),
    starts_with("Q18"),
    starts_with("Q19"),
    starts_with("Q20"),
    starts_with("Q21")
  )

#Identify metadata columns to keep at the front
edu_meta_cols <- c("ip_address", "progress", "duration_in_seconds",
               "finished", "recorded_date", "response_id", "group", "respondent_id")

#Identify and sort Q columns using mixedsort
edu_q_cols <- setdiff(colnames(Educator_Data_Clean), edu_meta_cols)
edu_q_cols_sorted <- mixedsort(edu_q_cols)

#Reorder the data frame
Educator_Data_Clean <- Educator_Data_Clean[, c(edu_meta_cols, edu_q_cols_sorted)]
```

# Introduction

Dr. Mariea Ross-Estrada, a faculty member at North Carolina State University's College of Veterinary Medicine, is exploring whether there are differences between the expectations of small animal primary care veterinary employers and veterinary educators regarding new graduates' competencies in dentistry. Through her own professional experience and conversations with colleagues, Dr. Ross-Estrada observed that many veterinarians must rely on on-the-job training to gain the skills necessary in small animal dentistry. These shared experiences prompted her to investigate whether there is a misalignment in what is taught in veterinary programs and what is expected in clinical practice.

To explore this question, Dr. Ross-Estrada distributed two surveys: one to medical directors and private practice owners and the other to primary care veterinary educators. Both surveys included similar questions regarding what early-career veterinarians are expected to have learned during their education and the skills they are expected to perform in practice.

This research has the potential to improve veterinary care by ensuring new graduates are better prepared to meet clinical demands. It could also support a smoother transition into practice for early-career veterinarians, reduce the burden of on-the-job training for employers, and promote overall workforce readiness in the field of small animal primary care.

## Research Question

How do small animal primary care employers (medical directors and practice owners) and primary care veterinary educators differ in regards to their expectations of early career veterinary graduates’ competencies in small animal dentistry?

## Statistical Questions

1.  Are there significant differences between educators and practice owners in their belief that new graduates are competent in key dental skills on their first day of practice?

2.  Is there a difference between educators and practice owners in their reports (educators’ actual teaching vs. owners’ perceptions) of which dental skills were taught in the pre-clinical DVM curriculum for recent graduates?

3.  Is there a difference between educators and practice owners in their level of agreement about whether specific dental skills should be taught pre-clinically?

4.  Do employers and educators differ in their expectations about how many dental procedures new graduates should complete during clinical training?

5.  Is there difference between the instructional formats in dentistry reported by DVM programs and the formats perceived by employers to have been completed by early career veterinarians?

6.  Do educators and employers differ in their views on which formats of clinical instruction in dentistry should be required for DVM students as part of their clinical training?

7.  Is there a difference between the clinical dentistry skills that educators report DVM students are learning during their clinical training and the skills that employers believe recent graduates have completed as part of their DVM program?

8.  Do educators and employers differ in their opinions about which clinical dentistry skills DVM students should be required to practice or learn during their clinical training?

# Data

## Data Description

Two separate surveys were administered to mutually exclusive groups: veterinary employers who have worked with students, and educators who have taught students. There was no overlap between these groups and they can be assumed to be independent.

The employer data set consists of responses from 29 participants answering 40 questions, while the educator data set includes 43 participants answering 34 questions. Each group was asked a single qualifying question to determine eligibility for participation, along with nine questions covering demographics and institutional context. Educators were then presented with 24 competency and sentiment-based questions, while employers answered 30 such items focused on professional expectations and training in veterinary medicine.

Survey questions took several forms. Some were binary (Yes/No), particularly those related to demographics and institutional affiliation. Others used a "select all that apply" format, commonly seen in questions asking respondents to identify procedures performed at their practice. Many of these questions were followed by Likert-scale items. The Likert scales were even-numbered and omitted a neutral option, which may have contributed to at least two instances where respondents selected both “agree” and “disagree” for the same item.

Several questions offered an “Other” response with a text box for elaboration. A few required numeric input, such as estimates of hours worked or the number of practicing veterinarians. These integer fields were not restricted by any upper bound, regardless of contextual reasonableness.

#### Global survey session metrics

```{r}
#| label: Data_Desc_01
#| fig-cap: "Assessing survey elapsed time distribution via box plots to understand engagement by survey group."
#| echo: false
#| fig-align: center

#Average time for survey completion
# Convert and clean both datasets
clean_employer <- Employer_Data[-c(1,2), ] %>%
  mutate(
    duration_in_seconds = as.numeric(duration_in_seconds),
    duration_minutes = duration_in_seconds / 60,
    group = "Employer"
  )

clean_educator <- Educator_Data[-c(1,2), ] %>%
  mutate(
    duration_in_seconds = as.numeric(duration_in_seconds),
    duration_minutes = duration_in_seconds / 60,
    group = "Educator"
  )

# Combine for optional single-geom plotting, or keep separate for layers
ggplot() +
  geom_boxplot(data = clean_employer,
               aes(x = group, y = duration_minutes),
               fill = "steelblue", alpha = 0.6) +
  geom_boxplot(data = clean_educator,
               aes(x = group, y = duration_minutes),
               fill = "darkorange", alpha = 0.6) +
  coord_cartesian(ylim = c(0, 80)) +  # Optional: zoom to 0–60 minutes
  labs(
    title = "Survey Completion Time by Group",
    x = "Respondent Group",
    y = "Duration (minutes)"
  ) +
  theme_minimal()
```

Survey completion time differed by group. Educators, on average, spent more time completing the survey than employers. While no follow-up question asked participants to explain their response time, this discrepancy may reflect greater engagement or a tendency for more elaborated responses among educators. It may also suggest a greater willingness among educators to participate more thoughtfully. The box plot above illustrates the distribution of survey duration (in minutes) by group.

```{r}
#| label: Data_Desc_02
#| fig-cap: "Assessing survey completion as a density curve to understand engagement by survey group."
#| echo: false
#| fig-align: center

#Average percent of survey completed
combined_progress <- bind_rows(
  Employer_Data[-c(1,2), ] %>% mutate(group = "Employer"),
  Educator_Data[-c(1,2), ] %>% mutate(group = "Educator")
) %>%
  mutate(progress = as.numeric(progress))

ggplot(combined_progress, aes(x = progress, fill = group)) +
  geom_density(alpha = 0.4) +
  labs(title = "Survey Progress Density by Group", x = "Progress (%)", y = "Density") +
  theme_minimal()

```

Regarding the proportion of the survey completed, employer responses were more variable—spanning the full range from partial to full completion. In contrast, educators tended to complete more of the survey, with a concentration near full completion and a less pronounced left tail. The density plot above visualizes these differences in survey progress across groups.

```{r}
#| label: Data_Desc_03
#| fig-cap: "Geographic distribution of survey respondents across the United States."
#| echo: false
#| fig-align: center

# Clean Employer Data
Employer_Clean <- Employer_Data %>%
  slice(-c(1,2)) %>%
  mutate(
    latitude = as.numeric(location_latitude),
    longitude = as.numeric(location_longitude)
  ) %>%
  filter(!is.na(latitude) & !is.na(longitude)) %>%
  mutate(group = "Employer")

# Clean Educator Data
Educator_Clean <- Educator_Data %>%
  slice(-c(1,2)) %>%
  mutate(
    latitude = as.numeric(location_latitude),
    longitude = as.numeric(location_longitude)
  ) %>%
  filter(!is.na(latitude) & !is.na(longitude)) %>%
  filter(
    between(latitude, 24, 50),
    between(longitude, -125, -65)
  ) %>%
  mutate(group = "Educator")

# Combine datasets
All_Locations <- bind_rows(Employer_Clean, Educator_Clean)

# Remove the specific outliers
All_Locations <- All_Locations %>%
  filter(!(
    (abs(latitude - 43.5204) < 1e-6 & abs(longitude + 80.2245) < 1e-6) |  
    (abs(latitude - 35.6339) < 1e-6 & abs(longitude - 140.0391) < 1e-6) | 
    (abs(latitude + 5.2521) < 1e-6 & abs(longitude - 39.7633) < 1e-6)     
  ))

# Get US map data
us_map <- map_data("state")

# Plot
ggplot() +
  geom_polygon(data = us_map, aes(x = long, y = lat, group = group), 
               fill = "gray90", color = "white") +
  geom_point(data = All_Locations, aes(x = longitude, y = latitude, color = group), 
             size = 2, alpha = 0.7) +
  coord_fixed(1.3) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ) +
  labs(title = "Geographic Distribution of Survey Respondents",
       color = "Group")
```

Since our analysis focused on the representativeness of the United States, we identified three sets of coordinates in the educator dataset corresponding to international institutions: the University of Guelph in Ontario, Canada; Zanzibar University in Chake, Tanzania; and Chiba University in Chiba, Japan. Employer survey participants were predominantly sampled from locations along the eastern seaboard of the United States, while educators were more widely distributed across the country. We highlight this to illustrate that perceptions of veterinary dental students—particularly among employers—may differ for individuals located far from the regions where most participants were sampled. Additionally, because many of the responses came from major metropolitan areas, our findings may underrepresent opinions regarding student knowledge gaps in more rural settings.

#### Educator metrics

```{r}
#| label: Data_Desc_04
#| fig-cap: "Educator contextualized background information about teaching and procedures."
#| echo: false
#| fig-align: center

# Remove metadata rows
Educator_Data_DESC03 <- Educator_Data[-c(1, 2), ]

# Ensure q27 is numeric and remove NA values
Educator_Data_DESC03$q27 <- as.numeric(Educator_Data_DESC03$q27)
Educator_Data_DESC03 <- Educator_Data_DESC03 %>% filter(!is.na(q27))

# Ensure q41_1 is numeric and remove NA values
Educator_Data_DESC03$q41_1 <- as.numeric(Educator_Data_DESC03$q41_1)
Educator_Data_DESC03 <- Educator_Data_DESC03 %>% filter(!is.na(q41_1) | !is.na(q27))

# Create ceiling-binned versions
Educator_Data_DESC03 <- Educator_Data_DESC03 %>%
  mutate(
    q27_binned = ceiling(q27),
    q41_1_binned = ceiling(q41_1)
  )

# Compute max y values for neat y-axis limits
max_count_p1 <- Educator_Data_DESC03 %>%
  count(q27_binned) %>%
  pull(n) %>%
  max(na.rm = TRUE)

max_count_p2 <- Educator_Data_DESC03 %>%
  count(q41_1_binned) %>%
  pull(n) %>%
  max(na.rm = TRUE)

# Plot 1: Years Teaching (with ceiling bins and clean y-axis)
p1 <- ggplot(Educator_Data_DESC03, aes(x = q27_binned)) +
  geom_bar(
    fill = "#4C72B0",
    color = "white"
  ) +
  scale_x_continuous(
    breaks = seq(0, max(Educator_Data_DESC03$q27_binned, na.rm = TRUE), by = 1)
  ) +
  scale_y_continuous(
    breaks = seq(0, max(max_count_p1 + 2, 2), by = 2)
  ) +
  labs(
    title = "Years Teaching DVM Students",
    x = "Years Teaching (Rounded Up)",
    y = "Number of Respondents"
  ) +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "grey85", linetype = "dashed"),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", size = 12),
    axis.title = element_text(size = 10)
  )

# Plot 2: Procedures per Week (with ceiling bins and clean y-axis)
p2 <- ggplot(Educator_Data_DESC03, aes(x = q41_1_binned)) +
  geom_bar(
    fill = "#DD8452",
    color = "white"
  ) +
  scale_x_continuous(
    breaks = seq(0, max(Educator_Data_DESC03$q41_1_binned, na.rm = TRUE), by = 1)
  ) +
  scale_y_continuous(
    breaks = seq(0, max(max_count_p2 + 2, 2), by = 2)
  ) +
  labs(
    title = "Dental Procedures per Week",
    x = "Procedures per Week (Rounded Up)",
    y = "Number of Educators"
  ) +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "grey85", linetype = "dashed"),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", size = 12),
    axis.title = element_text(size = 10)
  )

# Combine plots side by side
p1 + p2 + plot_layout(widths = c(1, 1))
```

Figure 3.4 provides an overview of the survey participants and their institutions. Most respondents reported having between 2 and 4 years of experience teaching veterinary students in clinical training. The distribution of years taught appears approximately normal, with fewer educators at the lower and upper ends of the experience range. Respondents also reported the number of dental procedures performed by their primary care service each week. While some outliers from busier institutions reported higher volumes, most educators estimated performing between 1 and 8 procedures per week.

#### Employer metrics

The employer dataset was considerably smaller than the educator dataset, which may reflect differences in motivation or willingness to complete the survey. Most employer respondents identified as practice owners, and approximately 75% indicated they were employed in private practice. This aligns with findings from the American Veterinary Medical Association (AVMA), which reported that "most veterinarians age 75 or younger, 83.9%, work in private clinical practice" [@avma2019].

```{r}
#| label: Data_Desc_06
#| fig-cap: ""
#| echo: false
#| warning: false
#| fig-align: center

#* Q33; Which of the following best describes your job setting or organization?
# Mapping vector
q33_labels <- c(
  "1" = "Independently owned single veterinary practice",
  "2" = "Independently owned group veterinary practice",
  "3" = "Group corporate veterinary practice",
  "4" = "Institution (e.g., aquarium, shelter, zoo, etc.)",
  "5" = "Industry/commercial",
  "6" = "Government (local, state, federal)",
  "7" = "Uniformed services",
  "8" = "Other"
)

# Clean and map responses, remove first two metadata rows
q33_clean <- Employer_Data[["q33"]] %>%
  .[-c(1,2)] %>%                          # Remove first two metadata rows
  na.omit() %>%
  as.character() %>%
  recode(!!!q33_labels)

# Create counts and percentages
q33_table <- as.data.frame(table(q33_clean)) %>%
  rename(`Job Setting` = q33_clean, `Count` = Freq) %>%
  mutate(Percentage = round(Count / sum(Count) * 100, 1))

# Pretty table
kable(q33_table, caption = "Job Setting or Organization: Counts and Percentages", align = "lrr", booktabs = TRUE)


#* Q37; Which of the following best describes the person who completed this survey?
# Mapping vector for Q37
q37_labels <- c(
  "1" = "Practice owner",
  "2" = "Practice manager/HR representative",
  "3" = "Associate veterinarian",
  "4" = "Other"
)

# Clean and map responses for Q37
q37_clean <- Employer_Data[["q37"]] %>%
  .[-c(1,2)] %>%             # Remove metadata rows
  na.omit() %>%
  as.character() %>%
  recode(!!!q37_labels)

# Create counts and percentages for Q37
q37_table <- as.data.frame(table(q37_clean)) %>%
  rename(`Respondent Role` = q37_clean, `Count` = Freq) %>%
  mutate(Percentage = round(Count / sum(Count) * 100, 1))

# Display table for Q37
kable(q37_table, caption = "Respondent Role: Counts and Percentages", align = "lrr", booktabs = TRUE)


```

Survey responses were also heavily skewed toward practice ownership, with 63.6% of employer participants identifying as owners. By comparison, the same AVMA survey reported that only 21.3% of veterinarians identified as practice owners. This discrepancy may be due to the smaller sample size or the possibility that practice owners were more willing to complete the survey than associates or other veterinary professionals.


```{r}
#| label: Data_Desc_07
#| fig-cap: "Average dental procedures employer survey participants indicated they performed each week."
#| echo: false
#| warning: false
#| fig-align: center
#* Q36; On average, how many dental procedures does your practice/oganization/institution perform each week


# Clean and prepare data
Employer_Clean <- Employer_Data %>%
  slice(-c(1,2)) %>%                       # remove metadata rows
  mutate(q36_1 = as.numeric(q36_1)) %>%
  filter(!is.na(q36_1)) %>%
  mutate(q36_1_binned = ceiling(q36_1))   # binning like p2

# Calculate max count for y-axis scaling
max_count_p36_1 <- Employer_Clean %>%
  count(q36_1_binned) %>%
  pull(n) %>%
  max(na.rm = TRUE)

# Plot similar to p2
p_q36_1 <- ggplot(Employer_Clean, aes(x = q36_1_binned)) +
  geom_bar(
    fill = "#DD8452",
    color = "white"
  ) +
  scale_x_continuous(
    breaks = seq(0, max(Employer_Clean$q36_1_binned, na.rm = TRUE), by = 1)
  ) +
  scale_y_continuous(
    breaks = seq(0, max(max_count_p36_1 + 2, 2), by = 2)
  ) +
  labs(
    title = "Individual Dental Procedures per Week",
    x = "Procedures per Week (Rounded Up)",
    y = "Number of Respondents"
  ) +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "grey85", linetype = "dashed"),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", size = 12),
    axis.title = element_text(size = 10)
  )

p_q36_1
```

Figure 3.5 shows that procedure counts tend to be lower in the employer group compared to the educator group, although the overall distribution shapes are similar. In both groups, the data exhibit a right-skewed pattern, with most respondents reporting lower procedure counts and a few outliers representing higher volumes. This pattern aligns with the intuitive understanding that while some practices or institutions have greater clinical demands on veterinarians, these cases are less common—at least based on the survey responses.

## Data Source

Survey data were collected using Qualtrics, a cloud-based experience management platform commonly used for gathering feedback and sentiment across workforce domains. Participants from the educator survey were recruited via email invitation sent by the researcher, using pre-existing contact lists. Dr. Ross-Estrada distributed the employer survey to her personal and professional networks online. Participation was voluntary and anonymous. There was no incentive offered for completing the survey.

## Preprocessing Description

Although the employer and educator data sets shared a similar structure, they were not identical. Most pre-processing steps were applied uniformly across both data sets, with minor deviations where needed.

The data sets were imported into the RStudio environment (version 2024.04.1 Build 748). A new variable was created to label the data source ("Educator" or "Employer") for later grouping and visualization. The existing respondent_id column served as a unique identifier and was treated as the primary key.

Initial cleaning involved removing extraneous metadata included by Qualtrics—such as survey start and end times, IP addresses, geolocation data, and question display logic—all of which were irrelevant to the analysis. These columns were trimmed to streamline the dataset for subsequent transformation and statistical work.

Column names in the original Qualtrics export were alphanumeric but often ambiguous and misleading. Many variable names did not match the corresponding survey question numbers. Our team manually mapped the exported column names to their corresponding survey questions and responses by referencing adjacent metadata fields and using deductive reasoning. This process allowed us to build an index-based column naming structure, which greatly improved the manageability and interpretability of the dataset.

Before diving into question-specific analysis, we first identified the subset of survey questions relevant to our research objectives. All unrelated or out-of-scope items were removed. This step reduced the employer dataset from 176 columns to 100, and the educator dataset from 171 columns to 102.

Several formatting inconsistencies also needed to be resolved. Some multi-select questions appeared in the form of comma-separated text responses within a single column, while others were exported into multiple binary columns. Additionally, for certain questions, a response option that received zero selections was dropped entirely by Qualtrics. To standardize these issues, we implemented a script to “explode” comma-separated responses into individual binary columns. For dropped columns, we manually reintroduced them as zero-filled dummy variables to preserve the full response structure.

Finally, we filtered out participants who answered less than half of the survey. We also excluded:

-   Employers who responded “No” to the question: “Do you work with early career veterinarians (someone who has graduated from a DVM program after May 2021)?”

-   Educators who responded “No” to: “Do you teach in any capacity of the dental curriculum at your institution?”

After all preprocessing steps, the final cleaned datasets consisted of 13 employer participants and 30 educator participants.

# Statistical Methods

#### Likert scale questions (statistical questions #1,3, 6, and 8)

These questions use a Likert scale with response options: Strongly Agree, Agree, Disagree, and Strongly Disagree across a variety of skills and procedures. Given the ordinal nature of these responses, we will use diverging stacked bar charts and summary tables to visually explore the data.  We will then use Mann-Whitney U Tests for formal hypothesis testing differences between groups.  Finally, a Ranked-Biserial Correlation (RBC) of effect size to find direction and magnitude of differences between groups.  

The Mann-Whitney U Test is a non-parametric test used to assess whether there is a statistically significant difference in the distributions of ordinal or continuous variables between two independent groups. In this case, educators and employers are assumed to be independent, consistent with the survey design.

This test is appropriate for Likert-scale data because:

* It does not assume normality (unlike the t-test),
* It respects the ordinal (ranked) structure of the data,
* It retains statistical power in small samples.

```{r}
#| label: mannwhitneytable
#| echo: false
#| warning: false
#| fig-cap: " "
#| fig-align: center

# Create a visual summary table
mw_data <- data.frame(
  Group = c("Group 1 (e.g., Educators)", "Group 2 (e.g., Employers)"),
  `Sample Size` = c("n₁", "n₂"),
  `Sum of Ranks` = c("R₁", "R₂"),
  `Mean Rank` = c("R₁ / n₁", "R₂ / n₂")
)

kable(mw_data, align = "lrrr", caption = "Mann-Whitney U Rank Summary Table", booktabs = TRUE)

```

In order to calculate the test statistic and p-value for a Mann-Whitney U Test:

* All observations are ranked together
* The sum of the ranks for each group is calculated ($R_1$ and $R_2$)
* The test statistic U is then calculated as the minimum of $U_1$ and $U_2$ where $U_1$ and $U_2$ are the following: $U_1 = R_1 - \frac{n_1(n_1 + 1)}{2},  U_2 = R_2 - \frac{n_2(n_2 + 1)}{2}$ 
* Then the p-value is calculated from the U distribution or normal approximation.

To complement the Mann-Whitney U tests, effect sizes were calculated using the rank-biserial correlation (RBC). While the p-value assesses whether a statistically significant difference exists between groups, the rank-biserial correlation quantifies the magnitude and direction of that difference—adding a layer of practical interpretation. An RBC value near 0 suggests little to no effect, while values near -1 or 1 suggest large differences favoring one group over the other.

Formally, the RBC is calculated directly from the Mann-Whitney U statistic:

$r_{rb} = \frac{2U}{n_1n_2} - 1$

Where:

* $U$ is the Mann-Whitney U statistic (typically the smaller of $U_1$ and $U_2$)  
* $n_1$ and $n_2$ are the sample sizes of the two independent groups  

This formula scales $U$ into the $[-1,1]$ interval, where:

* $r_{rb} = 0$ implies overlapping ranks (no difference)  
* $r_{rb} = 1$ implies every value in group 1 exceeds every value in group 2  
* $r_{rb} = -1$ implies every value in group 2 exceeds every value in group 1  


$U$ is derived by jointly ranking all observation across groups.  The sum or ranks ($R_1$ and $R_2$) are then used to calculate:

$U_1 = R_1 - \frac{n_1(n_1+1)}{2}$,  
$U_2 = R_2 - \frac{n_2(n_2+1)}{2}$

Then U = min($U_1$,$U_2$) and this value is substituted for $r_{rb}$.   

P-values and an RBC value will be applied to each statistical question listed above in order to formally analyze the data. 

#### Select all that apply questions (statistical questions #2, 5, and 7)

These questions asked the participants to “select all that apply” as it relates to skills in pre-clinical curriculum, format of dental instruction, and skills for clinical training respectively. We will utilize frequency tables and bar plots to explore the data for these questions and use Fisher’s Exact Test as a formal inference procedure for the comparison of the two groups. Dot plots will also be utilized for questions with many options to avoid crowding in bar graphs.

Fisher’s Exact Test works with categorical data with independent samples, in this case educators and employers. Based on the survey design from the client, it is again reasonable to assume that these two groups are independent. In this context, Fisher's Exact Test is preferred over Chi-Squared Tests due to the small sample size and therefore not meeting the expected count threshold that is required to proceed with Chi-Squared Tests.

```{r}
#| label: fishertable
#| echo: false
#| warning: false
#| fig-cap: " "
#| fig-align: center

# Create the generic table
table_data <- data.frame(
  Group = c("Group 1 (ex: Educators)", "Group 2 (ex: Employers)", "Column Totals"),
  `Outcome Present` = c("a", "c", "a + c"),
  `Outcome Absent` = c("b", "d", "b + d"),
  `Row Totals` = c("a + b", "c + d", "n = a + b + c + d"))
  
# Print the table
kable(table_data, align = "lrr", caption = "Fisher's Exact 2×2 Contingency Table", booktabs=TRUE)

```

Fisher's Exact Test operates under the null hypothesis that there is no association between the two variables. The alternative hypothesis is that there is an association. The p-value for the test is computed using the following formula:

$$p = \frac{(a+b)! \, (c+d)! \, (a+c)! \, (b+d)!}{a! \, b! \, c! \, d! \, n!}$$

P-values are then computed for each skill, format, etc. between the two groups (educators and employers) depending on the statistical question to be analyzed.


#### Numerical entry questions: (statistical question #4)

This question asks participants to enter a number related to the number of dental procedures that should be completed during training in different areas. We plan to produce box plots and/or histograms to visually examine the data. Depending on the normality or lack thereof of the distributions, we will then conduct either a two-sample t-test or a Mann-Whitney U Test. As mentioned above, the Mann Whitney U test is a non-parametric test that does not depend on the assumption of normality. If the assumption on normality is met, we can consider a two-sample t test for this analysis. However, from examination of the data, normality is not met and we will therefore utilize a Mann-Whitney U Test. The test follows the same method above as statistical questions #2, 5, and 7.

# Results

## Statistical Analysis

### S1 Are there significant differences between educators and practice owners in their belief that new graduates are competent in key dental skills on their first day of practice?


To assess whether there are differences in perceptions of new graduate competence in dentistry, educators and employers were asked a parallel question (Q4). Employers were asked to rate their expectation that early career veterinarians can competently perform 12 specific dental skills on their first day of employment. Educators were asked to rate whether they believe new graduates can perform those same skills competently. Responses were recorded on a 4-point Likert scale from 1 (Strongly Disagree) to 4 (Strongly Agree).


We used the Mann-Whitney U test to compare educator and employer responses for each skill, as this non-parametric test is well-suited for ordinal data from two independent groups. In addition to p-values, we calculated the ranked-biserial correlation (r_rb) to assess effect sizes and provide insight into the magnitude and direction of observed differences. Results are summarized with median ratings for each group, sample sizes, and effect size interpretations (e.g., small, medium, large). Significance stars were added to highlight results below conventional thresholds (e.g., p < 0.05), and results were sorted by p-value to aid interpretation.


```{r}
#| label: full_q4_mannwhitney
#| echo: false
#| warning: false
#| message: false
#| fig-align: center

library(dplyr)
library(tidyr)
library(purrr)
library(effectsize)
library(kableExtra)

# Add group identifiers
educator_cleaned <- Educator_Data_Clean %>%
  mutate(group = "Educator")

employer_cleaned <- Employer_Data_Clean %>%
  mutate(group = "Employer")

# Define Q4 variable names
q4_vars <- paste0("Q04_", sprintf("%02d", 1:12))

# Subset and combine
combined_data <- bind_rows(
  educator_cleaned %>% select(all_of(q4_vars)) %>% mutate(group = "Educator"),
  employer_cleaned %>% select(all_of(q4_vars)) %>% mutate(group = "Employer")
)

# Reshape and clean
q4_long <- combined_data %>%
  pivot_longer(cols = all_of(q4_vars), names_to = "Skill", values_to = "Rating_num") %>%
  mutate(Rating_num = as.numeric(Rating_num)) %>%
  filter(!is.na(Rating_num))

# Run tests and format
mw_results_q4 <- q4_long %>%
  group_by(Skill) %>%
  filter(n_distinct(group) == 2) %>%
  summarise(
    p_value = wilcox.test(Rating_num ~ group, exact = FALSE)$p.value,
    r_rb = effectsize::rank_biserial(Rating_num ~ group)$r_rank_biserial,
    median_educator = median(Rating_num[group == "Educator"]),
    median_employer = median(Rating_num[group == "Employer"]),
    n_educator = sum(group == "Educator"),
    n_employer = sum(group == "Employer"),
    .groups = "drop"
  ) %>%
mutate(
  stars = cut(p_value,
              breaks = c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
              labels = c("***", "**", "*", ".", "")),
  `P-Value` = ifelse(is.na(p_value), "N/A",
                     ifelse(stars != "", paste0(sprintf("%.3f", p_value), " ", stars),
                            sprintf("%.3f", p_value))),
  `Med (Edu)` = sprintf("%.0f", median_educator),
  `Med (Emp)` = sprintf("%.0f", median_employer),
  `N (Edu / Emp)` = paste0(n_educator, " / ", n_employer),
  r_rb_num = as.numeric(r_rb),
  r_rb_fmt = ifelse(is.na(r_rb_num), "N/A", sprintf("%.2f", r_rb_num)),
  Interpretation = case_when(
    is.na(r_rb_num) ~ "N/A (Insufficient Data)",
    abs(r_rb_num) >= 0.50 ~ paste0("Large ", ifelse(r_rb_num > 0, "positive", "negative"), " effect"),
    abs(r_rb_num) >= 0.30 ~ paste0("Medium ", ifelse(r_rb_num > 0, "positive", "negative"), " effect"),
    abs(r_rb_num) >= 0.10 ~ paste0("Small ", ifelse(r_rb_num > 0, "positive", "negative"), " effect"),
    TRUE ~ "Negligible effect"
  )
) %>%
select(
  Skill,
  `Med (Edu)`,
  `Med (Emp)`,
  `N (Edu / Emp)`,
  `P-Value`,
  `r_rb` = r_rb_fmt,
  Interpretation
)


# Human-readable skill names
skill_labels <- c(
  "Q04_01" = "Periodontal probing",
  "Q04_02" = "Dental radiography - positioning",
  "Q04_03" = "Dental radiography - interpretation",
  "Q04_04" = "Scaling",
  "Q04_05" = "Polishing",
  "Q04_06" = "Simple extractions - canine",
  "Q04_07" = "Simple extractions - feline",
  "Q04_08" = "Open extractions - canine (single root)",
  "Q04_09" = "Open extractions - feline (single root)",
  "Q04_10" = "Open extractions - canine (multi-root)",
  "Q04_11" = "Open extractions - feline (multi-root)",
  "Q04_12" = "Fluoride treatment"
)

mw_results_q4$Skill <- recode(mw_results_q4$Skill, !!!skill_labels)

# Compact table formatting
kable(mw_results_q4,
      format = "latex",
      booktabs = TRUE,
      caption = "Mann-Whitney U-Test Results for Competency Ratings",
      align = "c",
      row.names = FALSE) %>%
  kable_styling(latex_options = c("striped", "hold_position", "scale_down"))

```

Overall, educators and employers had generally similar perceptions about new graduate competence in most dental skills, with no statistically significant differences observed at the conventional p < 0.05 threshold. However, some notable trends emerged. Dental radiography, interpretation showed a borderline significant result (p = 0.052) with a medium positive effect size (r_rb = 0.36), suggesting that educators may be slightly more confident than employers in new graduates’ ability to interpret dental radiographs. For most other skills, effect sizes were negligible or small, and p-values were well above 0.05, indicating no meaningful differences in perceived competence between groups. In a few cases, such as open extractions, canine (multi-root), employers had slightly higher confidence than educators (small  negative effect size), but again, these differences were not statistically significant. The fluoride treatment item had a large effect size, but due to the very small sample (2 educators, 1 employer), it is not meaningful or generalizable.



### S2 Is there a difference between educators and practice owners in their reports (educators’ actual teaching vs. owners’ perceptions) of which dental skills were taught in the pre-clinical DVM curriculum for recent graduates?


To evaluate whether differences exist between DVM educators and employers in their understanding of which dental skills are taught in the pre-clinical curriculum, we analyzed responses to parallel survey questions. Educators were asked to indicate which of seven core dentistry skills are taught as part of their pre-clinical courses (Q12), while employers were asked which skills they believe recent graduates were taught prior to clinical training (Q16). Since the two groups answered different question numbers about the same underlying skills, we first aligned the datasets by renaming employer variables to match educator labels. This harmonization allowed for direct comparison of responses skill-by-skill across the two groups.

After cleaning and reshaping the data into long format, we filtered out invalid or missing entries and conducted Fisher’s Exact Tests for each skill. This non-parametric test is appropriate for evaluating categorical (yes/no) outcomes in small samples, especially when comparing proportions between two independent groups. Only skills for which both groups provided non-missing responses were included in the final analysis. 



```{r}
#| label: question_2prep
#| echo: false
#| warning: false
#| message: false
#| include: false
#| fig-align: center

# 1. Define skill name lookup table
skill_lookup <- tibble::tibble(
  Skill_Format = c("Q12_01", "Q12_02", "Q12_03", "Q12_04", "Q12_05", "Q12_06", "Q12_07"),
  Skill = c(
    "Oral examination and charting",
    "Dental cleaning (scaling and polishing)",
    "Dental radiograph acquisition",
    "Dental radiograph interpretation",
    "Simple extraction technique",
    "Local nerve blocks",
    "Instrument identification and handling"
  )
)

# 2. Prepare Employer data (rename Q16_01–07 to Q12_01–07)
employer_q16 <- Employer_Data_Clean %>%
  select(Q16_01:Q16_07) %>%
  rename_with(~ paste0("Q12_", sprintf("%02d", 1:7)), .cols = everything()) %>%
  mutate(group = "Employer")

# 3. Prepare Educator data
educator_q12 <- Educator_Data_Clean %>%
  select(Q12_01:Q12_07) %>%
  mutate(group = "Educator")

# 4. Combine both datasets
combined_q12_q16 <- bind_rows(educator_q12, employer_q16)

# 5. Pivot to long format for analysis
combined_long <- combined_q12_q16 %>%
  mutate(across(starts_with("Q12_"), as.character)) %>%
  pivot_longer(
    cols = starts_with("Q12_"),
    names_to = "Skill_Format",
    values_to = "Taught_raw"
  ) %>%
  mutate(
    Taught = na_if(Taught_raw, "N/A"),
    Taught = na_if(Taught, ""),
    Taught = as.numeric(Taught)
  ) %>%
  filter(!is.na(Taught), group %in% c("Educator", "Employer"))

# 6. Run Fisher’s exact tests
test_results <- combined_long %>%
  group_by(Skill_Format) %>%
  summarise(
    tab = list(table(group, Taught)),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    dims = list(dim(tab)),
    is_valid = !is.null(dims[[1]]) && all(dims[[1]] >= 2)
  ) %>%
  filter(is_valid) %>%
  mutate(
    p_value = tryCatch(fisher.test(tab)$p.value, error = function(e) NA_real_),
    n_educator = if ("Educator" %in% rownames(tab)) sum(tab["Educator", ]) else NA_integer_,
    n_employer = if ("Employer" %in% rownames(tab)) sum(tab["Employer", ]) else NA_integer_,
    prop_educator = if ("Educator" %in% rownames(tab) && "1" %in% colnames(tab)) {
      tab["Educator", "1"] / sum(tab["Educator", ])
    } else NA_real_,
    prop_employer = if ("Employer" %in% rownames(tab) && "1" %in% colnames(tab)) {
      tab["Employer", "1"] / sum(tab["Employer", ])
    } else NA_real_
  ) %>%
  ungroup() %>%
  select(-tab, -is_valid, -dims) %>%
  left_join(skill_lookup, by = "Skill_Format") %>%
  select(Skill, p_value, n_educator, n_employer, prop_educator, prop_employer) %>%
  arrange(p_value)

```

```{r}
#| label: question_2
#| echo: false
#| warning: false
#| fig-align: center

kable(test_results,
      digits = 3,
      caption = "Fisher's Exact Test Results by Skill",
      format = "markdown")  # or "latex"/"html" based on output

```

Overall, educators and employers showed general agreement about which dental skills are taught in the pre-clinical DVM curriculum, with no statistically significant differences found across any of the seven skills analyzed. Proportions of reported instruction were similar between groups, with slight variations. For example, more educators than employers reported that dental radiograph acquisition and simple extraction techniques are taught, while more employers believed that oral examination and charting and dental cleaning are included. However, none of these differences reached significance, and all p-values were well above 0.05. These findings suggest a shared but imperfect understanding between groups, with minor discrepancies that are not statistically meaningful.



### S3 Is there a difference between educators and practice owners in their level of agreement about whether specific dental skills should be taught pre-clinically?

To assess whether educators and employers differ in their beliefs about which dental skills should be included in the pre-clinical DVM curriculum, we analyzed responses to parallel Likert-scale questions: Q13 (educators) and Q17 (employers). Both groups rated the importance of teaching 12 specific dentistry skills using a 4-point Likert scale ranging from 1 (Strongly Disagree) to 4 (Strongly Agree). 

Because the data were ordinal and responses came from two independent groups, the Mann-Whitney U test was used to compare educator and employer ratings for each skill individually. This non-parametric test is appropriate for comparing ordinal data between two independent groups, especially when the assumptions for parametric tests may not hold. Ratings were reshaped into long format and analyzed skill-by-skill. The results table presents medians for each group, sample sizes, p-values with significance indicators, effect size estimates (rank-biserial correlation), and an interpretation of the magnitude and direction of group differences. 




```{r question_3prep, include=FALSE}
#| label: question_3prep
#| include: false
#| echo: false
#| warning: false
#| message: false

library(dplyr)
library(tidyr)
library(purrr)
library(effectsize)
library(kableExtra)

# Skill name lookup
skill_labels_q3 <- c(
  "Skill_01" = "Periodontal probing",
  "Skill_02" = "Dental radiography - positioning",
  "Skill_03" = "Dental radiography - interpretation",
  "Skill_04" = "Scaling",
  "Skill_05" = "Polishing",
  "Skill_06" = "Simple extractions - canine",
  "Skill_07" = "Simple extractions - feline",
  "Skill_08" = "Open extractions - canine (single root)",
  "Skill_09" = "Open extractions - feline (single root)",
  "Skill_10" = "Open extractions - canine (multi-root)",
  "Skill_11" = "Open extractions - feline (multi-root)",
  "Skill_12" = "Fluoride treatment"
)

# Select and format data
q13_vars <- paste0("Q13_", sprintf("%02d", 1:12))  # Educators
q17_vars <- paste0("Q17_", sprintf("%02d", 1:12))  # Employers

educator_q13 <- Educator_Data_Clean %>%
  select(all_of(q13_vars)) %>%
  rename_with(~ paste0("Skill_", sprintf("%02d", 1:12)), .cols = everything()) %>%
  mutate(group = "Educator")

employer_q17 <- Employer_Data_Clean %>%
  select(all_of(q17_vars)) %>%
  rename_with(~ paste0("Skill_", sprintf("%02d", 1:12)), .cols = everything()) %>%
  mutate(group = "Employer")

combined_q3 <- bind_rows(educator_q13, employer_q17)

q3_long <- combined_q3 %>%
  pivot_longer(
    cols = starts_with("Skill_"),
    names_to = "Skill",
    values_to = "Rating_num"
  ) %>%
  mutate(Rating_num = as.numeric(Rating_num)) %>%
  filter(!is.na(Rating_num))

mw_results_q3 <- q3_long %>%
  group_by(Skill) %>%
  filter(n_distinct(group) == 2) %>%
  summarise(
    p_value = wilcox.test(Rating_num ~ group, exact = FALSE)$p.value,
    r_rb = effectsize::rank_biserial(Rating_num ~ group)$r_rank_biserial,
    median_educator = median(Rating_num[group == "Educator"]),
    median_employer = median(Rating_num[group == "Employer"]),
    n_educator = sum(group == "Educator"),
    n_employer = sum(group == "Employer"),
    .groups = "drop"
  ) %>%
  mutate(
    stars = cut(p_value,
                breaks = c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
                labels = c("***", "**", "*", ".", "")),
    P_Value = ifelse(is.na(p_value), "N/A",
                     ifelse(stars != "", paste0(sprintf("%.3f", p_value), " ", stars),
                            sprintf("%.3f", p_value))),
    `Med (Edu)` = sprintf("%.0f", median_educator),
    `Med (Emp)` = sprintf("%.0f", median_employer),
    `N (Edu / Emp)` = paste0(n_educator, " / ", n_employer),
    r_rb_fmt = sprintf("%.2f", r_rb),
    Interpretation = case_when(
      is.na(r_rb) ~ "N/A (Insufficient Data)",
      abs(r_rb) >= 0.50 ~ paste0("Large ", ifelse(r_rb > 0, "positive", "negative"), " effect"),
      abs(r_rb) >= 0.30 ~ paste0("Medium ", ifelse(r_rb > 0, "positive", "negative"), " effect"),
      abs(r_rb) >= 0.10 ~ paste0("Small ", ifelse(r_rb > 0, "positive", "negative"), " effect"),
      TRUE ~ "Negligible effect"
    )
  ) %>%
  arrange(p_value) %>%
  mutate(Skill = recode(Skill, !!!skill_labels_q3)) %>%
  select(
    Skill,
    `Med (Edu)`,
    `Med (Emp)`,
    `N (Edu / Emp)`,
    `P-Value` = P_Value,
    `r_rb` = r_rb_fmt,
    Interpretation
  )

invisible(mw_results_q3)

```


```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-align: center

kable(mw_results_q3,
      caption = "Mann-Whitney U-Test Results for Pre-Clinical Dentistry Training",
      align = "c",
      row.names = FALSE,
      escape = FALSE,
      latex_options = c("scale_down", "HOLD_position", "striped")) %>%
  add_header_above(c(" " = 1,
                     "Median" = 2,
                     "N" = 1,
                     " " = 1,
                     "Effect Size" = 1,
                     " " = 1),
                   line = FALSE) %>%
  column_spec(1, width = "1.4in") %>%
  column_spec(2, width = "0.5in") %>%
  column_spec(3, width = "0.5in") %>%
  column_spec(4, width = "0.7in") %>%
  column_spec(5, width = "0.65in") %>%
  column_spec(6, width = "0.65in") %>%
  column_spec(7, width = "1.1in") %>%
  kable_styling(font_size = 9, full_width = FALSE, position = "center", latex_options = "scale_down") %>%
  row_spec(0, bold = TRUE)

```

Educators and employers generally agreed on the importance of teaching core dental skills in the pre-clinical DVM curriculum, with most skills receiving high median ratings (4 = Strongly Agree) from both groups. However, statistically significant differences were observed for two skills. Open extractions skill, feline (multi-root) showed a significant difference (p = 0.002) with a large negative effect size (r_rb = -0.60), indicating employers placed greater importance on teaching this skill than educators. Fluoride treatment also differed significantly (p = 0.013) with a very large negative effect size (r_rb = -1.00), though the very small sample of educators (n = 3) limits interpretation.

Several other extraction-related skills showed small negative effect sizes and borderline p-values, suggesting a trend toward employers assigning slightly more importance to teaching advanced extractions pre-clinically, though these differences did not reach statistical significance. For all other skills, differences were negligible or small, and p-values were well above 0.05, indicating broad consensus on their pre-clinical importance.



### S4 Do employers and educators differ in their expectations about how many dental procedures new graduates should complete during clinical training?
To assess whether educators and practice owners differ in their expectations for the number of dental procedures students should complete during clinical training, we compared responses from Question 19 of the educator survey and Question 24 of the employer survey. Respondents indicated the number of procedures they believe should be completed in four clinical settings: primary care, dentistry rotations, shelter medicine, and lab animal medicine. 

Before selecting an appropriate statistical test, we assessed normality using the Shapiro-Wilk test, which showed that the data were not normally distributed in any group-setting combination (all p-values < 0.05). Given the non-normal distribution and the ordinal or skewed nature of the data, we used the Mann-Whitney U test—a non-parametric alternative to the independent samples t-test—to compare the distributions of responses between educators and employers for each clinical setting. This test is appropriate when comparing two independent groups without assuming normality or equal variances.

```{r}
#| label: question_4prep
#| include: false
#| echo: false
#| warning: false
#| message: false

library(dplyr)
library(tidyr)
library(purrr)
library(effectsize)
library(gt)

# 1. Define variables
educator_vars <- paste0("Q19_0", 1:4)
employer_vars <- paste0("Q24_0", 1:4)
setting_labels <- c("Primary_Care", "Dentistry_Rotation", "Shelter_Medicine", "Lab_Animal")

# 2. Select and rename columns
educator_proc <- Educator_Data_Clean %>%
  select(all_of(educator_vars)) %>%
  mutate(group = "Educator")
employer_proc <- Employer_Data_Clean %>%
  select(all_of(employer_vars)) %>%
  mutate(group = "Employer")

colnames(educator_proc)[1:4] <- setting_labels
colnames(employer_proc)[1:4] <- setting_labels

# 3. Combine and reshape
combined_proc <- bind_rows(educator_proc, employer_proc) %>%
  pivot_longer(
    cols = all_of(setting_labels),
    names_to = "Setting",
    values_to = "Procedure_Count"
  ) %>%
  filter(!is.na(Procedure_Count)) %>%
  mutate(Procedure_Count = as.numeric(Procedure_Count))

# 4. Calculate Mann-Whitney and effect size
mw_results <- combined_proc %>%
  group_by(Setting) %>%
  summarise(
    p_value = wilcox.test(Procedure_Count ~ group)$p.value,
    r_rb = effectsize::rank_biserial(Procedure_Count ~ group)$r_rank_biserial,
    median_educator = median(Procedure_Count[group == "Educator"]),
    median_employer = median(Procedure_Count[group == "Employer"]),
    n_educator = sum(group == "Educator"),
    n_employer = sum(group == "Employer"),
    .groups = "drop"
  ) %>%
  mutate(
    stars = cut(p_value,
                breaks = c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
                labels = c("***", "**", "*", ".", "")),
    P_Value = ifelse(is.na(p_value), "N/A",
                     ifelse(stars != "", paste0(sprintf("%.3f", p_value), " ", stars),
                            sprintf("%.3f", p_value))),
    r_rb_fmt = sprintf("%.2f", r_rb),
    Interpretation = case_when(
      is.na(r_rb) ~ "N/A (Insufficient Data)",
      abs(r_rb) >= 0.50 ~ paste0("Large ", ifelse(r_rb > 0, "positive", "negative"), " effect"),
      abs(r_rb) >= 0.30 ~ paste0("Medium ", ifelse(r_rb > 0, "positive", "negative"), " effect"),
      abs(r_rb) >= 0.10 ~ paste0("Small ", ifelse(r_rb > 0, "positive", "negative"), " effect"),
      TRUE ~ "Negligible effect"
    )
  ) %>%
  arrange(p_value) %>%
  select(
    Setting,
    median_educator,
    median_employer,
    n_educator,
    n_employer,
    P_Value,
    r_rb_fmt,
    Interpretation
  )
```

```{r}
#| echo: false
#| label: q4graph
#| warning: false
#| message: false
#| fig-align: center


mw_results %>%
  gt() %>%
  tab_header(
    title = md("**Mann-Whitney U Test: Expected Number of Dental Procedures**"),
    subtitle = "Comparison of Educator and Employer Expectations by Clinical Setting"
  ) %>%
  cols_label(
    Setting = "Setting",
    median_educator = "Med (Edu)",
    median_employer = "Med (Emp)",
    n_educator = "N (Edu)",
    n_employer = "N (Emp)",
    P_Value = "P-Value",
    r_rb_fmt = "r_rb",
    Interpretation = "Effect Size"
  ) %>%
  cols_align(align = "center") %>%
  fmt_number(columns = vars(median_educator, median_employer), decimals = 1) %>%
  fmt_number(columns = vars(P_Value), decimals = 3) %>%
  fmt_number(columns = vars(r_rb_fmt), decimals = 2) %>%
  tab_options(
    data_row.padding = px(1),
    table.font.size = px(10),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(11),
    column_labels.font.size = px(11),
    source_notes.font.size = px(9)
  ) %>%
  opt_align_table_header("center")
```


Educators and employers showed broad agreement on the number of dental procedures that students should complete during clinical training, with no statistically significant differences across the four clinical settings evaluated. However, a notable trend was observed in the dentistry rotation, where employers had higher expectations than educators (median = 6.5 vs. 4.0), with a borderline p-value (p = 0.080) and a medium negative effect size (r_rb = -0.39), suggesting that employers may expect more hands-on experience in this setting.

In the remaining settings—lab animal medicine, shelter medicine, and primary care—differences in medians were small, and effect sizes were negligible to small, with p-values well above 0.05, indicating no meaningful disagreement between groups. These findings suggest that while general expectations are aligned, employers may place slightly greater emphasis on procedural experience in dedicated dentistry rotations.



### S5 Is there difference between the instructional formats in dentistry reported by DVM programs and the formats perceived by employers to have been completed by early career veterinarians?

Both educators and employers were asked a question relating to the format of instruction during the clinical year.

On question 20 of their survey, employers were asked: "What format of clinical instruction in dentistry do you believe that the early career veterinarians (individuals who have graduated from a DVM program after May 2021) hired into your practice/organization/institution completed as part of their DVM training? Select all that apply."

On question 16 of their survey, educators were asked: "What format of instruction in dentistry does your DVM program provide during the clinical year? Select all that apply."

Since the question was of the format "select all that apply," participants were able to select more than one response and percentages will not add to 100%. Results reported below are the percentages of their respective group (educators or employers) that selected the given format of instruction. Percentages were utilized due to the difference in sample size.

Educators and employers were also given the option to enter their own responses to the question. One employer opted to write in "Cadaver." Two educators entered their own responses with one saying "rounds" and the other saying "topic seminars do occur with some rotations during case rounds when dental cases are chosen to present."

Educators and employers differed in their perceptions of clinical dental instruction formats completed by early career veterinarians. While a majority in both groups acknowledged didactic instruction, educators reported higher rates of wet lab (75% vs. 46.2%) and live patient training (96.4% vs. 38.5%) compared to employers. Conversely, employers more frequently identified simulation training (30.8% vs. 14.3%) than educators.

```{r}
#| label: formatgraph
#| echo: false
#| warning: false
#| fig-cap: "Perceived Clinical Instruction Formats in Dentistry Completed by Early Career Veterinarians"
#| fig-align: center
#Add custom labels to final bar graph
option_labels <- c(
  "01" = "Didactic",
  "02" = "Simulation",
  "03" = "Wet Lab",
  "04" = "Live Patient",
  "05" = "None of these"
)

# Count total respondents per group
n_employers <- nrow(Employer_Data_Clean)
n_educators <- nrow(Educator_Data_Clean)

#Employer data: calculate percentage of total dataset
employer_percent <- Employer_Data_Clean %>%
  select(Q20_01:Q20_05) %>%
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "option", values_to = "count") %>%
  mutate(group = "Employer",
         option = str_replace(option, "Q20_", ""),
         percent = count / n_employers * 100)

#Educator data: calculate percentage of total dataset
educator_percent <- Educator_Data_Clean %>%
  select(Q16_01:Q16_05) %>%
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "option", values_to = "count") %>%
  mutate(group = "Educator",
         option = str_replace(option, "Q16_", ""),
         percent = count / n_educators * 100)

#Combine
combined_percent <- bind_rows(employer_percent, educator_percent) %>%
  mutate(option = case_when(
    option == "01" ~ "Didactic",
    option == "02" ~ "Simulation",
    option == "03" ~ "Wet Lab",
    option == "04" ~ "Live Patient",
    option == "05" ~ "None of These",
    TRUE ~ option
  ))

#Keep the order of the bars as the order of the question
combined_percent$option <- factor(combined_percent$option,
                                  levels = c("Didactic", "Simulation", "Wet Lab", "Live Patient", "None of These"))

#Plot percentages
ggplot(combined_percent, aes(x = option, y = percent, fill = group)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title ="Dentistry Formats",
       x = "Format of Instruction",
       y = "Percent of Respondents",
       fill = "Group") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  ylim(0, max(combined_percent$percent) + 10) +
  geom_text(aes(label = paste0(round(percent, 1), "%")), 
          position = position_dodge(width = 0.9), 
          vjust = -0.3, size = 3)
```

Figure 5.1 shows the percentages of each format selected and Table 5.4 shows the p-value associated with performing Fisher's Exact Test on each format.

Didactic instruction was selected by 57.1% of educators and 61.5% of employers, with no statistically significant difference between the groups (p=1). There was also no statistically significant difference in simulation (p=0.237) and wet lab instruction (p=0.0885).

On the other hand, live patient instruction was selected by 96.4% of educators, but only 38.5% of employers. This difference was found to be statistically significant with a p-value of 0.000105.

"None of these" was also found to be statistically significant at the 5% level. No educators selected that none of these formats were believed to have been used, but 23.1% of employers selected that none of the formats were believed to be utilized in the clinical year.

```{r}
#| label: question_5b
#| echo: false
#| warning: false
#| fig-cap: " "
#| fig-align: center
# Total respondents per group
n_employers <- nrow(Employer_Data_Clean)
n_educators <- nrow(Educator_Data_Clean)

# Function to get counts of selected responses
get_counts <- function(data, vars, n_total) {
  data %>%
    select(all_of(vars)) %>%
    summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) %>%
    pivot_longer(everything(), names_to = "option", values_to = "count") %>%
    mutate(percent = count / n_total * 100,
           option = str_replace(option, "Q[0-9]+_", ""))
}

# Get counts for each group
employer_counts <- get_counts(Employer_Data_Clean, paste0("Q20_0", 1:5), n_employers)
educator_counts <- get_counts(Educator_Data_Clean, paste0("Q16_0", 1:5), n_educators)

# Merge counts by option
combined_counts <- educator_counts %>%
  rename(educator_count = count, educator_percent = percent) %>%
  inner_join(
    employer_counts %>% rename(employer_count = count, employer_percent = percent),
    by = "option"
  )

# Map option codes to labels
combined_counts <- combined_counts %>%
  mutate(Format = case_when(
    option == "01" ~ "Didactic",
    option == "02" ~ "Simulation",
    option == "03" ~ "Wet Lab",
    option == "04" ~ "Live Patient",
    option == "05" ~ "None of These"
  ))

# Calculate p-values using Fisher's Exact Test for each option
combined_counts <- combined_counts %>%
  rowwise() %>%
  mutate(
    p_value = {
      a <- educator_count
      b <- n_educators - a
      c <- employer_count
      d <- n_employers - c
      contingency_matrix <- matrix(c(a, b, c, d), nrow = 2, byrow = TRUE)
      fisher.test(contingency_matrix)$p.value
    }
  ) %>%
  ungroup()

# Select and arrange columns for output
final_table <- combined_counts %>%
  select(Format, educator_percent, employer_percent, p_value) %>%
  arrange(match(Format, c("Didactic", "Simulation", "Wet Lab", "Live Patient", "None of These")))


#Format percentages nicely
final_table <- final_table %>%
  mutate(
    educator_percent = round(educator_percent, 1),
    employer_percent = round(employer_percent, 1),
    p_value = signif(p_value, 3)
  ) %>%
  rename(
    `% of Educators Selected` = educator_percent,
    `% of Employers Selected` = employer_percent,
    `P-value` = p_value
  )

# Add significance stars
final_table <- final_table %>%
  mutate(
    Significance = case_when(
      `P-value` < 0.001 ~ "***",
      `P-value` < 0.01  ~ "**",
      `P-value` < 0.05  ~ "*",
      TRUE              ~ ""
    ),
    `P-value` = paste0(`P-value`, " ", Significance)
  ) %>%
  select(-Significance)  # Optional: remove extra column if not needed

kable(final_table, caption = "Perceived Formats of Clinical Dental Instruction in DVM Programs", align = "lrr", booktabs=TRUE)

```

### S6 Do educators and employers differ in their views on which formats of clinical instruction in dentistry should be required for DVM students as part of their clinical training?

In Question #21 of the employer survey, participants were asked:

• “Which of the following types of clinical instruction in dentistry do you think DVM students should be required to complete as part of a DVM program? Select one response for each of the instructional types listed below.”

The analogous question for educators appeared as Question #17 and stated:

• “Which of the following types of instruction do you think DVM students should be required to complete as part of their clinical training? Select one response for each type of instruction listed below.”

This question aimed to assess perceptions regarding which types of dental instruction should be required in the veterinary medical curriculum across both survey groups. Table 5.3 presents the Mann–Whitney U-test results, summary statistics, effect sizes, and ranked-biserial correlation outputs.


```{r}
#| label: 6a_MannWhit_Table
#| echo: false
#| warning: false
#| message: false
#| include: true
#| fig-align: center

# Define the matching question columns
educator_cols <- paste0("Q17_", sprintf("%02d", 1:7))
employer_cols <- paste0("Q21_", sprintf("%02d", 1:7))

# Create human-friendly labels (ensure 'Other - Text' is distinct)
instruction_labels <- c(
"Q17_01" = "Didactic", "Q21_01" = "Didactic",
"Q17_02" = "Simulation", "Q21_02" = "Simulation",
"Q17_03" = "Wet Lab",  "Q21_03" = "Wet Lab",
"Q17_04" = "Live Patient", "Q21_04" = "Live Patient",
"Q17_05" = "None of these", "Q21_05" = "None of these",
"Q17_06" = "Other",  "Q21_06" = "Other",
"Q17_07" = "Other - Text (Specify)" # Clarified label
)

# Initialize a data frame to store all results
all_mw_results_df <- tibble()

for (i in seq_along(educator_cols)) {
q_edu <- educator_cols[i]
q_emp <- employer_cols[i]

# Extract values and convert to numeric, ensuring consistency
edu_data <- Educator_Data_Clean |> # Using base R pipe |> (R 4.1+)
 select(Response = all_of(q_edu)) |>
 mutate(Response = as.numeric(Response), Group = "Educator") |>
 filter(!is.na(Response))

emp_data <- Employer_Data_Clean |>
 select(Response = all_of(q_emp)) |>
 mutate(Response = as.numeric(Response), Group = "Employer") |>
 filter(!is.na(Response))

# Combine data for the current skill
combined_skill_data <- bind_rows(edu_data, emp_data)

# Ensure both groups have at least 2 observations to run test
if (n_distinct(combined_skill_data$Group) == 2 &&
 sum(combined_skill_data$Group == "Educator") >= 2 &&
 sum(combined_skill_data$Group == "Employer") >= 2) {

 # Ensure 'Group' is a factor with desired order for consistent sign of r_rb
 # A positive r_rb means the group listed second (Educator) tends to have higher values.
 combined_skill_data$Group <- factor(combined_skill_data$Group, levels = c("Employer", "Educator"))

 # Perform Wilcoxon test (result kept for p-value)
 test_result <- wilcox.test(Response ~ Group, data = combined_skill_data, exact = FALSE)

 # Calculate ranked-biserial correlation using rank_biserial() with the formula and data
 r_rb_object <- effectsize::rank_biserial(Response ~ Group, data = combined_skill_data)
 r_rb <- r_rb_object$r_rank_biserial # Extract the specific value

 # Calculate medians
 med_edu <- median(edu_data$Response, na.rm = TRUE)
 med_emp <- median(emp_data$Response, na.rm = TRUE)

 # Store results
 current_result <- tibble(
 Skill_ID = q_edu,
 p_value = test_result$p.value,
 r_rb = r_rb,
 median_educator = med_edu,
 median_employer = med_emp,
 n_educators = nrow(edu_data),
 n_employers = nrow(emp_data)
 )
 all_mw_results_df <- bind_rows(all_mw_results_df, current_result)
} else {
 # Handle cases where a test can't be run (e.g., insufficient data in one group)
 current_result <- tibble(
 Skill_ID = q_edu,
 p_value = NA, # Assign NA for p-value if test not run
 r_rb = NA,  # Assign NA for r_rb if test not run
 median_educator = median(edu_data$Response, na.rm = TRUE),
 median_employer = median(emp_data$Response, na.rm = TRUE),
 n_educators = nrow(edu_data),
 n_employers = nrow(emp_data)
 )
 all_mw_results_df <- bind_rows(all_mw_results_df, current_result)
}
}

# Add significance stars and readable labels for final table
final_results_table <- all_mw_results_df |>
mutate(
 Skill = recode(Skill_ID, !!!instruction_labels),
 stars = cut(p_value,
    breaks = c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
    labels = c("***", "**", "*", ".", "")),
 P_Value = ifelse(is.na(p_value), "N/A", ifelse(stars != "", paste0(sprintf("%.3f", p_value), " ", stars), sprintf("%.3f", p_value))),
 # Median values to be printed with no float (integers)
 `Median (Educator)` = sprintf("%.0f", median_educator),
 `Median (Employer)` = sprintf("%.0f", median_employer),
 `n (Educator / Employer)` = paste0(n_educators, " / ", n_employers),
 `Ranked-Biserial Correlation (r_rb)` = sprintf("%.2f", r_rb),
 Interpretation = case_when(
 is.na(r_rb) ~ "N/A (Insufficient Data)",
 abs(r_rb) >= 0.50 ~ paste0("Large ", ifelse(r_rb > 0, "positive", "negative"), " effect"),
 abs(r_rb) >= 0.30 ~ paste0("Medium ", ifelse(r_rb > 0, "positive", "negative"), " effect"),
 abs(r_rb) >= 0.10 ~ paste0("Small ", ifelse(r_rb > 0, "positive", "negative"), " effect"),
 TRUE ~ "Negligible effect"
 )
) |>
  filter(P_Value != "N/A") |>
select(
 Skill,
 `Med (Edu)` = `Median (Educator)`, # Concise header
 `Med (Emp)` = `Median (Employer)`, # Concise header
 `N (Edu / Emp)` = `n (Educator / Employer)`, # Concise header
 `P-Value` = P_Value, # Consistent capitalization, remove space for LaTeX
 `r_rb` = `Ranked-Biserial Correlation (r_rb)`, # Concise header
 Interpretation
) |>
arrange(`P-Value`) # Arrange by the new column name

# Render the final table with kableExtra for advanced formatting
kable(final_results_table,
   caption = "Mann-Whitney U-Test Results for Clinical Instruction Formats",
   align = "c",
   row.names = FALSE,
   escape = FALSE, # Allows LaTeX in headers and content
   latex_options = c("scale_down", "HOLD_position", "striped") # latex_options belongs INSIDE kable()
   ) |>
 add_header_above(c(" " = 1,
          "Median" = 2,
          "N" = 1,
          " " = 1,
          "Effect Size" = 1,
          " " = 1),
         line = FALSE) |>
 # Significantly reduce column widths to tighten the table.
 # Adjusted widths for a tighter fit.
 column_spec(1, width = "1.0in") |>
 column_spec(2, width = "0.5in") |>
 column_spec(3, width = "0.5in") |>
 column_spec(4, width = "0.65in") |>
 column_spec(5, width = "0.55in") |>
 column_spec(6, width = "0.75in") |>
 column_spec(7, width = "1.1in")
```


The Mann–Whitney U-test revealed a statistically significant difference ($p = 0.004$) between employers and educators when evaluating the importance of Live Patients. Although both groups reported a median rating of 4, suggesting similar central tendencies, the ranked-biserial correlation ($r_{rb} = -0.35$) indicates a medium negative effect. This suggests that employers generally rated Live Patient instruction as more essential than educators did. The difference in ranking distributions is statistically meaningful.

No statistically significant differences were found between the groups regarding the instructional types Simulation ($p = 0.366$), Didactic ($p = 0.467$), or Wet Lab ($p = 0.658$). The ranked-biserial correlations for Simulation ($r_{rb} = 0.18$) and Didactic ($r_{rb} = 0.13$) were small and positive, indicating a slight tendency for educators to rate these types of instruction as more important than employers. The effect size for Wet Lab ($r_{rb} = 0.05$) was negligible. None of these differences reached statistical significance.

The instructional categories “None of These” and “Other” were excluded from Table 5.3 due to insufficient data for statistical testing.

```{r}
#| label: s6_diverging_stacked_bar
#| echo: false
#| warning: false
#| message: false
#| include: true
#| fig-height: 8 # Adjust height as needed for readability
#| fig-width: 10 # Adjust width as needed for readability
#| fig-cap: "Diverging Stacked Bar Chart of Perceived Importance/Competence of Clinical Instruction Formats by Group (Educator vs. Employer)."

library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats) # For fct_rev and fct_relevel
library(scales)  # For scales::label_percent()

# Define the matching question columns (from your S6 code)
educator_cols <- paste0("Q17_", sprintf("%02d", 1:7))
employer_cols <- paste0("Q21_", sprintf("%02d", 1:7))

# Create human-friendly labels (from your S6 code)
instruction_labels <- c(
  "Q17_01" = "Didactic", "Q21_01" = "Didactic",
  "Q17_02" = "Simulation", "Q21_02" = "Simulation",
  "Q17_03" = "Wet Lab", "Q21_03" = "Wet Lab",
  "Q17_04" = "Live Patient", "Q21_04" = "Live Patient",
  "Q17_05" = "None of these", "Q21_05" = "None of these",
  "Q17_06" = "Other", "Q21_06" = "Other",
  "Q17_07" = "Other - Text (Specify)" # Clarified label
)

# --- 1. Prepare and Combine Data ---
# Select and rename educator data
edu_long <- Educator_Data_Clean |>
  select(all_of(educator_cols)) |>
  mutate(Group = "Educator", ID = row_number()) |>
  pivot_longer(
    cols = -c(Group, ID),
    names_to = "Question_ID",
    values_to = "Response"
  ) |>
  filter(Question_ID %in% c("Q17_01", "Q17_02", "Q17_03", "Q17_04"))

# Select and rename employer data
emp_long <- Employer_Data_Clean |>
  select(all_of(employer_cols)) |>
  mutate(Group = "Employer", ID = row_number()) |>
  pivot_longer(
    cols = -c(Group, ID),
    names_to = "Question_ID",
    values_to = "Response"
  ) |>
  filter(Question_ID %in% c("Q21_01", "Q21_02", "Q21_03", "Q21_04"))

# Combine both datasets
combined_data_s6 <- bind_rows(edu_long, emp_long) |>
  filter(!is.na(Response)) # Remove NA responses

# --- 2. Create a unified "Skill" label for plotting and Likert factor ---
full_label_map <- instruction_labels[c("Q17_01", "Q17_02", "Q17_03", "Q17_04")]
names(full_label_map) <- c("Q17_01", "Q17_02", "Q17_03", "Q17_04")
full_label_map <- c(full_label_map,
                    "Q21_01" = "Didactic", "Q21_02" = "Simulation",
                    "Q21_03" = "Wet Lab", "Q21_04" = "Live Patient")

combined_data_s6 <- combined_data_s6 |>
  mutate(
    Skill = recode(Question_ID, !!!full_label_map),
    Response_Factor = factor(Response,
                             levels = 1:4,
                             labels = c("1 - Strongly Disagree", "2 - Disagree", "3 - Agree", "4 - Strongly Agree"),
                             ordered = TRUE)
  )

# --- 3. Calculate Proportions and Ns for plotting ---
total_n_s6 <- combined_data_s6 |>
  group_by(Skill, Group) |>
  summarise(N = n(), .groups = 'drop')

plot_data_s6 <- combined_data_s6 |>
  group_by(Skill, Group, Response_Factor) |>
  summarise(Count = n(), .groups = 'drop_last') |>
  mutate(Proportion = Count / sum(Count)) |>
  ungroup() |>
  # Reversing the divergence logic: Educator is now negative, Employer is positive
  mutate(Proportion_Diverge = ifelse(Group == "Educator", -Proportion, Proportion))

plot_data_s6 <- plot_data_s6 |>
  left_join(total_n_s6, by = c("Skill", "Group"))

# --- 4. Create the Diverging Stacked Bar Chart (Mirror Style) ---

# === REORDER SKILLS HERE (TOP TO BOTTOM) ===
custom_skill_order <- c("Live Patient", "Simulation", "Didactic", "Wet Lab")
plot_data_s6$Skill <- factor(plot_data_s6$Skill, levels = rev(custom_skill_order)) # Use rev() to display top-to-bottom

likert_colors <- c(
  "1 - Strongly Disagree" = "#de2d26",
  "2 - Disagree" = "#fc9272",
  "3 - Agree" = "#a1d99b",
  "4 - Strongly Agree" = "#2ca25f"
)

ggplot(plot_data_s6, aes(y = Skill)) +
  # === EDUCATOR BARS ON LEFT (NEGATIVE X) ===
  geom_bar(data = plot_data_s6 |> filter(Group == "Educator"),
           aes(x = Proportion_Diverge, fill = Response_Factor),
           stat = "identity", width = 0.8) +
  # === EMPLOYER BARS ON RIGHT (POSITIVE X) ===
  geom_bar(data = plot_data_s6 |> filter(Group == "Employer"),
           aes(x = Proportion_Diverge, fill = Response_Factor),
           stat = "identity", width = 0.8) +

  geom_vline(xintercept = 0, color = "grey50", linetype = "solid", size = 0.5) +

  # Update N label positioning for new group layout
  geom_text(data = total_n_s6 |> filter(Group == "Educator"),
            aes(x = -0.05, y = Skill, label = paste0("N=", N)), # Position left of center for Educator
            hjust = 1, size = 3, color = "black") +
  geom_text(data = total_n_s6 |> filter(Group == "Employer"),
            aes(x = 0.05, y = Skill, label = paste0("N=", N)), # Position right of center for Employer
            hjust = 0, size = 3, color = "black") +

  scale_x_continuous(
    breaks = seq(-1, 1, 0.1),
    labels = function(x) {
      paste0(abs(x * 100), "%")
    },
    limits = c(-1.1, 1.1)
  ) +
  scale_fill_manual(
    values = likert_colors,
    drop = FALSE,
    name = "Likert Response",
    guide = guide_legend(reverse = FALSE)
  ) +
  labs(
    title = "Perceived Importance/Competence of Clinical Instruction Formats",
    subtitle = "Distribution of Likert Scale Responses by Group (Educator vs. Employer)",
    x = "Proportion of Responses",
    y = "Clinical Instruction" # === UPDATED Y-AXIS LABEL ===
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.grid.major.y = element_blank(),
    legend.title = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold")
  ) +
  # Update group labels for new layout
  annotate("text", x = -0.75, y = Inf, label = "Educator", # Educator label on left
           vjust = 1.5, hjust = 0.5, size = 4, fontface = "bold", color = "black") +
  annotate("text", x = 0.75, y = Inf, label = "Employer", # Employer label on right
           vjust = 1.5, hjust = 0.5, size = 4, fontface = "bold", color = "black")
```

Figure 5.2 presents the proportional distribution of responses for each group. While a moderate difference was observed between groups regarding the use of live patients, the majority of participants in both groups selected Strongly Agree, and notably, no participants selected Disagree or Strongly Disagree—indicating strong consensus that live patient experience should be required in clinical training.

The response distributions for simulation and wet lab instruction appear nearly identical visually, and statistical analysis confirmed no significant differences between groups for these skills. Interestingly, the distribution for didactic instruction looks somewhat different from those two, despite also showing no statistically significant difference.

This highlights an important point: even when one bar chart appears to show more variability in proportions than another, the overall ranking of responses between groups may still be similar. The Mann-Whitney U-test is designed to detect consistent directional differences in how groups rank responses. If those patterns are weak, inconsistent, or cancel each other out, the test will indicate no significant difference—even when the visualizations suggest otherwise. Differences in sample size and variability can also exaggerate or smooth over apparent contrasts that don’t reflect meaningful differences in central tendency.

### S7 Is there a difference between the clinical dentistry skills that educators report DVM students are learning during their clinical training and the skills that employers believe recent graduates have completed as part of their DVM program?

Both educators and employers were asked a question related to skills learned and practiced during the clinical year.

On question 25 of their survey, employers were asked: "Which of the following skills do you think that individuals who graduated with a DVM degree after May 2021 completed during the clinical training portion of their DVM program? Select all that apply."

On question 20 of their survey, educators were asked: "Which of the following skills are DVM students at your institution practicing/learning during the clinical training portion of the DVM program? Select all that apply."

Since the question was of the format "select all that apply," participants were able to select more than one response and percentages will not add to 100%. Results reported below are the percentages of their respective group (educators or employers) that selected the given format of instruction. Percentages were utilized due to the difference in sample size.

Educators and employers were also given the option to enter their own responses to the question. No employers entered any text responses. Six educators entered text responses of the following:

-   "OvaVet gel application, Crown amputation, Sealant application for UCF, oral tumor biopsy and excision, orthodontia"
-   "Nerve blocks, barrier sealant and bonded sealant application, jaw fracture repair, root canals"
-   "Often- bonded sealants; Sometimes- root canals, restorations, jaw fracture repair"
-   "Oral biopsy, root planning, bonded sealants"
-   "extractions only if performed on that patient"
-   "Not all students see all types of extractions"

```{r}
#| label: question_7a
#| echo: false
#| warning: false
#| fig-cap: "Clinical Training Skills Perceived by Educators vs Employers in DVM Programs"
#| fig-align: center
#| fig.width: 10
#| fig.height: 7
# Labels for each skill
option_labels <- c(
  "01" = "Oral Exam/Charting",
  "02" = "Radiographic Positioning",
  "03" = "Radiographic Interpretation",
  "04" = "Scaling",
  "05" = "Polishing",
  "06" = "Canine Closed Extraction",
  "07" = "Feline Closed Extraction",
  "08" = "Canine Open Extraction - Single Root",
  "09" = "Feline Open Extraction - Single Root",
  "10" = "Canine Open Extraction - Canines/Multiple Roots",
  "11" = "Feline Open Extraction - Canines/Multiple Roots",
  "12" = "Fluoride Treatment",
  "13" = "None of These"
)

# Respondent counts
n_employers <- nrow(Employer_Data_Clean)
n_educators <- nrow(Educator_Data_Clean)

# Valid variable names (exclude Q20_14 or Q20_NA)
valid_options <- sprintf("%02d", 1:13)
educator_vars <- paste0("Q20_", valid_options)
employer_vars <- paste0("Q25_", valid_options)

# Function to compute percent selected
get_counts <- function(data, vars, n_total) {
  data %>%
    select(all_of(vars)) %>%
    summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) %>%
    pivot_longer(cols = everything(), names_to = "option", values_to = "count") %>%
    mutate(
      percent = count / n_total * 100,
      option = str_replace(option, "Q[0-9]+_", "")
    )
}

# Get percentages for each group
educator_percent <- get_counts(Educator_Data_Clean, educator_vars, n_educators) %>%
  mutate(Group = "Educators")

employer_percent <- get_counts(Employer_Data_Clean, employer_vars, n_employers) %>%
  mutate(Group = "Employers")

# Combine and label
combined_percent <- bind_rows(educator_percent, employer_percent) %>%
  filter(option %in% names(option_labels)) %>%
  mutate(
    Label = factor(option_labels[option], levels = rev(option_labels)),
    Group = factor(Group, levels = c("Educators", "Employers"))
  )

# Dot plot
ggplot(combined_percent, aes(x = percent, y = Label, color = Group)) +
  geom_point(aes(size = percent), alpha = 0.8) +
  geom_text(aes(label = paste0(round(percent, 1), "%")), hjust = -0.2, color = "black", fontface = "bold") +
  scale_color_manual(values = c("Educators" = "#1f78b4","Employers" = "#e31a1c"))+
  scale_size(range = c(3, 8), guide = "none") +
  labs(
    x = "% of respondents in respective groups",
    y = NULL,
    title = "Clinical Training Skills",
    subtitle = "Each dot shows percent of Eduactors or Employers who selected each skill",
    color = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "top",
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold"),
    axis.text.y = element_text(size = 10),
    plot.title.position = "plot"
  ) +
  xlim(0, max(combined_percent$percent, na.rm = TRUE) + 10)

```

```{r}
#| label: question_7b
#| echo: false
#| warning: false
#| fig-cap: " "
#| fig-align: center
# Define labels for each of the 13 skills
option_labels <- c(
  "01" = "Oral Exam/Charting",
  "02" = "Radiographic Positioning",
  "03" = "Radiographic Interpretation",
  "04" = "Scaling",
  "05" = "Polishing",
  "06" = "Canine Closed Extraction",
  "07" = "Feline Closed Extraction",
  "08" = "Canine Open Extraction - Single Root",
  "09" = "Feline Open Extraction - Single Root",
  "10" = "Canine Open Extraction - Canines/Multiple Roots",
  "11" = "Feline Open Extraction - Canines/Multiple Roots",
  "12" = "Fluoride Treatment",
  "13" = "None of These"
)

# Total respondents per group
n_employers <- nrow(Employer_Data_Clean)
n_educators <- nrow(Educator_Data_Clean)

# Generic function to summarize counts from 1s
get_counts <- function(data, vars, n_total) {
  data %>%
    select(all_of(vars)) %>%
    summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) %>%
    pivot_longer(everything(), names_to = "option", values_to = "count") %>%
    mutate(
      percent = count / n_total * 100,
      option = str_replace(option, "Q[0-9]+_", "")  # Remove question prefix
    )
}

# Get counts for employers (Q25_01 to Q25_13)
employer_counts <- get_counts(Employer_Data_Clean, paste0("Q25_", str_pad(1:13, 2, pad = "0")), n_employers)

# Get counts for educators (Q20_01 to Q20_13)
educator_counts <- get_counts(Educator_Data_Clean, paste0("Q20_", str_pad(1:13, 2, pad = "0")), n_educators)

# Merge the two groups by skill option
combined_counts <- educator_counts %>%
  rename(educator_count = count, educator_percent = percent) %>%
  inner_join(
    employer_counts %>% rename(employer_count = count, employer_percent = percent),
    by = "option"
  )

# Add skill labels
combined_counts <- combined_counts %>%
  mutate(Skill = option_labels[option])

# Run Fisher’s Exact Test for each skill
combined_counts <- combined_counts %>%
  rowwise() %>%
  mutate(
    p_value = {
      a <- educator_count
      b <- n_educators - a
      c <- employer_count
      d <- n_employers - c
      fisher.test(matrix(c(a, b, c, d), nrow = 2, byrow = TRUE))$p.value
    }
  ) %>%
  ungroup()

# Format the final table
final_table <- combined_counts %>%
  select(Skill, educator_percent, employer_percent, p_value) %>%
  mutate(
    educator_percent = round(educator_percent, 1),
    employer_percent = round(employer_percent, 1),
    p_value = signif(p_value, 3)
  ) %>%
  rename(
    `% of Educators` = educator_percent,
    `% of Employers` = employer_percent,
    `P-value` = p_value
  ) %>%
  arrange(match(Skill, option_labels))

# Add significance stars
final_table <- final_table %>%
  mutate(
    Significance = case_when(
      `P-value` < 0.001 ~ "***",
      `P-value` < 0.01  ~ "**",
      `P-value` < 0.05  ~ "*",
      TRUE              ~ ""
    ),
    `P-value` = paste0(`P-value`, " ", Significance)
  ) %>%
  select(-Significance)  # Optional: remove extra column if not needed


# Print a clean table
kable(final_table, caption = "Perceived Skills Learned of Clinical Dental Instruction in DVM Programs", align = "lrr", booktabs=TRUE)
```

Figure 5.3 shows the percentage of employers and educators who selected the given skills and Table 5.6 shows the p-values associated with Fisher's Exact Test and their significance.

For eleven of the thirteen skills, educators reported a higher percentage of selection. The two skills for which employers reported higher perceived learning were feline open extraction - canines/multiple roots and fluoride treatment.

At the 5% significance level, the difference in skill selection between employers and educators was found to be significant for seven skills:

-   Radiographic Positioning (p=0.0284)
-   Scaling (p=0.0084)
-   Polishing (p=0.00331)
-   Canine Closed Extraction (p=0.0102)
-   Feline Closed Extraction (p=0.0102)
-   Canine Open Extraction - Single Root (p=0.0102)
-   Feline Open Extraction - Single Root (p=0.0193)
-   Canine Open Extraction - Canines/Multiple Roots (p=0.0193)

For all skills with statistically significant differences, educators reported higher percentages than employers.

The largest difference was found in polishing. Educators selected this skill 82.1% of the time, while employers selected this skill only 30.8% of the time. The resulting p-value was 0.00331.

### S8 Do educators and employers differ in their opinions about which clinical dentistry skills DVM students should be required to practice or learn during their clinical training?

In question #26 of the employers’ version of the survey, participants were asked:

* “Which of the following skills do you think that DVM students should be required to practice/learn as part of the clinical training portion of a DVM program? Select one response for each of the skills listed below.”

The corresponding question for educators was survey question #21.

This question aimed to assess opinions on which clinical dentistry skills should be required in DVM student training. Table 5.5 presents the Mann–Whitney U-test results, summary statistics, effect sizes, and ranked-biserial correlation outputs.


```{r question_8a}
#| label: 8a_MannWhit_Table
#| echo: false
#| warning: false
#| message: false
#| include: true
#| fig-align: center


# Define the matching question columns
educator_cols <- paste0("Q21_", sprintf("%02d", 1:13))
employer_cols <- paste0("Q26_", sprintf("%02d", 1:13))

# Create human-friendly labels
procedure_labels_map <- c(
  "Q21_01" = "Oral exam/charting",
  "Q21_02" = "Radiographic positioning",
  "Q21_03" = "Radiographic interpretation",
  "Q21_04" = "Scaling",
  "Q21_05" = "Polishing",
  "Q21_06" = "Canine closed extraction",
  "Q21_07" = "Feline closed extraction",
  "Q21_08" = "Canine open extraction - Single root",
  "Q21_09" = "Feline open extraction - Single root",
  "Q21_10" = "Canine open extraction - Canines/multiple roots",
  "Q21_11" = "Feline open extraction - Canines/multiple roots",
  "Q21_12" = "Fluoride treatment",
  "Q21_13" = "None"
)

# Initialize a data frame to store all results
all_mw_results_df_s8 <- tibble()

for (i in seq_along(educator_cols)) {
  q_edu <- educator_cols[i]
  q_emp <- employer_cols[i]

  # Extract values and convert to numeric, ensuring consistency
  edu_data <- Educator_Data_Clean %>%
    select(Response = all_of(q_edu)) %>%
    mutate(Response = as.numeric(Response), Group = "Educator") %>%
    filter(!is.na(Response))

  emp_data <- Employer_Data_Clean %>%
    select(Response = all_of(q_emp)) %>%
    mutate(Response = as.numeric(Response), Group = "Employer") %>%
    filter(!is.na(Response))

  # Combine data for the current skill
  combined_skill_data <- bind_rows(edu_data, emp_data)

  # Ensure both groups have at least 2 observations to run test
  if (n_distinct(combined_skill_data$Group) == 2 &&
      sum(combined_skill_data$Group == "Educator") >= 2 &&
      sum(combined_skill_data$Group == "Employer") >= 2) {

    # Ensure 'Group' is a factor with desired order for consistent sign of r_rb
    # A positive r_rb means the group listed second (Educator) tends to have higher values.
    combined_skill_data$Group <- factor(combined_skill_data$Group, levels = c("Employer", "Educator"))

    # Perform Wilcoxon test (result kept for p-value)
    # Using exact=FALSE for larger N as done in previous example
    test_result <- wilcox.test(Response ~ Group, data = combined_skill_data, exact = FALSE)

    # Calculate ranked-biserial correlation using rank_biserial()
    r_rb_object <- effectsize::rank_biserial(Response ~ Group, data = combined_skill_data)
    r_rb <- r_rb_object$r_rank_biserial # Extract the specific value

    # Calculate medians
    med_edu <- median(edu_data$Response, na.rm = TRUE)
    med_emp <- median(emp_data$Response, na.rm = TRUE)

    # Store results
    current_result <- tibble(
      Skill_ID = q_edu,
      p_value = test_result$p.value,
      r_rb = r_rb,
      median_educator = med_edu,
      median_employer = med_emp,
      n_educators = nrow(edu_data),
      n_employers = nrow(emp_data)
    )
    all_mw_results_df_s8 <- bind_rows(all_mw_results_df_s8, current_result)
  } else {
    # Handle cases where a test can't be run (e.g., insufficient data in one group)
    current_result <- tibble(
      Skill_ID = q_edu,
      p_value = NA, # Assign NA for p-value if test not run
      r_rb = NA,    # Assign NA for r_rb if test not run
      median_educator = median(edu_data$Response, na.rm = TRUE),
      median_employer = median(emp_data$Response, na.rm = TRUE),
      n_educators = nrow(edu_data),
      n_employers = nrow(emp_data)
    )
    all_mw_results_df_s8 <- bind_rows(all_mw_results_df_s8, current_result)
  }
}

# Add significance stars and readable labels for final table
final_results_table_s8 <- all_mw_results_df_s8 %>%
  mutate(
    Skill = recode(Skill_ID, !!!procedure_labels_map), # Use the new labels map
    stars = cut(p_value,
                breaks = c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
                labels = c("***", "**", "*", ".", "")),
    P_Value = ifelse(is.na(p_value), "N/A", ifelse(stars != "", paste0(sprintf("%.3f", p_value), " ", stars), sprintf("%.3f", p_value))),
    # Median values to be printed with no float (integers)
    `Median (Educator)` = sprintf("%.0f", median_educator),
    `Median (Employer)` = sprintf("%.0f", median_employer),
    `n (Educator / Employer)` = paste0(n_educators, " / ", n_employers),
    `Ranked-Biserial Correlation (r_rb)` = sprintf("%.2f", r_rb),
    Interpretation = case_when(
      is.na(r_rb) ~ "N/A (Insufficient Data)",
      abs(r_rb) >= 0.50 ~ paste0("Large ", ifelse(r_rb > 0, "positive", "negative"), " effect"),
      abs(r_rb) >= 0.30 ~ paste0("Medium ", ifelse(r_rb > 0, "positive", "negative"), " effect"),
      abs(r_rb) >= 0.10 ~ paste0("Small ", ifelse(r_rb > 0, "positive", "negative"), " effect"),
      TRUE ~ "Negligible effect"
    )
  ) %>%
    # Filter out rows where P_Value is "N/A"
    filter(P_Value != "N/A") %>%
  # Rename columns directly in select() for kableExtra
  select(
    Skill,
    `Med (Edu)` = `Median (Educator)`, # Concise header
    `Med (Emp)` = `Median (Employer)`, # Concise header
    `N (Edu / Emp)` = `n (Educator / Employer)`, # Concise header
    `P-Value` = P_Value, # Consistent capitalization, remove space for LaTeX
    `r_rb` = `Ranked-Biserial Correlation (r_rb)`, # Concise header
    Interpretation
  ) %>%
  arrange(`P-Value`) # Arrange by the new column name

# Render the final table with kableExtra for advanced formatting
# Adjust column headers for better display and tightness
kable(final_results_table_s8,
      caption = "Mann-Whitney U-Test Results for Clinical Procedures",
      align = "c",
      row.names = FALSE,
      escape = FALSE, # Allows LaTeX in headers and content
      latex_options = c("scale_down", "HOLD_position", "striped") # latex_options belongs INSIDE kable()
      ) %>%
  # Use add_header_above for multi-line headers, make them more concise
  add_header_above(c(" " = 1,
                     "Median" = 2,
                     "N" = 1,
                     " " = 1,
                     "Effect Size" = 1,
                     " " = 1),
                   line = FALSE) %>%
  # Significantly reduce column widths to tighten the table.
  # Adjusted widths for a tighter fit.
  column_spec(1, width = "1.3in") %>%  # Skill - might need more space for longer procedure names
  column_spec(2, width = "0.5in") %>% # Med (Edu)
  column_spec(3, width = "0.5in") %>% # Med (Emp)
  column_spec(4, width = "0.65in") %>%  # N (Edu / Emp)
  column_spec(5, width = "0.55in") %>%  # P-Value
  column_spec(6, width = "0.75in") %>% # r_rb
  column_spec(7, width = "1.1in")   # Interpretation
```

Nearly every skill reported a median of 4 with exception to the Feline Open Extraction - Canines/Multiple Roots whose central tendency was 3.  Note that this exception is the only row that report fewer responses than the others.  Small but consistent differences were found for several clinical procedures, with educators tending to rate certain skills as more important than employers.

Specifically, educators rated the following procedures as more important (higher median values) compared to employers:

- Feline closed extraction (effect size: $r = 0.27$)
- Canine open extraction – single root ($r = 0.27$)
- Feline open extraction – single root ($r = 0.23$)
- Canine open extraction – canines/multiple roots ($r = 0.23$)
- Feline open extraction – canines/multiple roots ($r = 0.26$)
- Radiographic interpretation ($r = 0.12$)
- Canine closed extraction ($r = 0.11$)

All of these showed small positive effect sizes, indicating that educators placed slightly more emphasis on these procedures than employers. However, none of these differences reached conventional levels of statistical significance (i.e., $p < 0.05$), though some were marginal ($p ≈ 0.06–0.09$).  Negligible differences were found for the following skills:

- Radiographic positioning ($r = -0.10$)
- Scaling ($r = -0.09$)
- Polishing ($r = -0.03$)
- Oral examination/charting ($r = 0.03$)

These results suggest that educators and employers generally agreed on the importance of these particular skills, with only very small or no measurable differences between groups.

```{r}
#| label: s8_diverging_stacked_bar_final
#| echo: false
#| warning: false
#| message: false
#| include: true
#| fig-height: 10
#| fig-width: 12
#| fig-cap: "Diverging Stacked Bar Chart of Perceived Competence of Clinical Procedures by Group (Educator vs. Employer)."

# Required libraries
library(dplyr)
library(tidyr)
library(ggplot2)

# --- (STEP 1) Define ALL necessary objects ---

# Define the question columns from your S8 analysis
educator_cols_s8 <- paste0("Q21_", sprintf("%02d", 1:12))
employer_cols_s8 <- paste0("Q26_", sprintf("%02d", 1:12))

# Define the human-readable labels for the procedures
procedure_labels_s8 <- c(
  "Q21_01" = "Oral exam/charting", "Q26_01" = "Oral exam/charting",
  "Q21_02" = "Radiographic positioning", "Q26_02" = "Radiographic positioning",
  "Q21_03" = "Radiographic interpretation", "Q26_03" = "Radiographic interpretation",
  "Q21_04" = "Scaling", "Q26_04" = "Scaling",
  "Q21_05" = "Polishing", "Q26_05" = "Polishing",
  "Q21_06" = "Canine closed extraction", "Q26_06" = "Canine closed extraction",
  "Q21_07" = "Feline closed extraction", "Q26_07" = "Feline closed extraction",
  "Q21_08" = "Canine open extraction - Single root", "Q26_08" = "Canine open extraction - Single root",
  "Q21_09" = "Feline open extraction - Single root", "Q26_09" = "Feline open extraction - Single root",
  "Q21_10" = "Canine open extraction - Canines/multiple roots", "Q26_10" = "Canine open extraction - Canines/multiple roots",
  "Q21_11" = "Feline open extraction - Canines/multiple roots", "Q26_11" = "Feline open extraction - Canines/multiple roots",
  "Q21_12" = "Fluoride treatment", "Q26_12" = "Fluoride treatment"
)

# Pivot educator data from wide to long format
edu_long_s8 <- Educator_Data_Clean |>
  select(all_of(educator_cols_s8)) |>
  mutate(Group = "Educator", ID = row_number()) |>
  pivot_longer(cols = -c(Group, ID), names_to = "Question_ID", values_to = "Response") |>
  # Filter out responses that contain a comma (e.g., "1,4")
  filter(!grepl(",", Response))

# Pivot employer data from wide to long format
emp_long_s8 <- Employer_Data_Clean |>
  select(all_of(employer_cols_s8)) |>
  mutate(Group = "Employer", ID = row_number()) |>
  pivot_longer(cols = -c(Group, ID), names_to = "Question_ID", values_to = "Response")


# --- (STEP 2) Filter data and calculate proportions ---

# Combine, filter, and create factors
combined_data_s8_filtered <- bind_rows(edu_long_s8, emp_long_s8) |>
  filter(!is.na(Response)) |>
  # Ensure Response is numeric before converting to factor, as it might be character due to filtering
  mutate(Response = as.numeric(Response)) |>
  mutate(Skill = recode(Question_ID, !!!procedure_labels_s8)) |>
  filter(Skill != "Fluoride treatment") |>
  mutate(
    Response_Factor = factor(Response,
                             levels = 1:4,
                             labels = c("1 - Strongly Disagree", "2 - Disagree", "3 - Agree", "4 - Strongly Agree"),
                             ordered = TRUE)
  )

# Calculate total responses (N) for labeling
total_n_s8 <- combined_data_s8_filtered |>
  group_by(Skill, Group) |>
  summarise(N = n(), .groups = 'drop')

# Calculate proportions for the stacked bars
plot_data_s8 <- combined_data_s8_filtered |>
  group_by(Skill, Group, Response_Factor) |>
  summarise(Count = n(), .groups = 'drop_last') |>
  mutate(Proportion = Count / sum(Count)) |>
  ungroup() |>
  mutate(Proportion_Diverge = ifelse(Group == "Educator", -Proportion, Proportion)) |>
  left_join(total_n_s8, by = c("Skill", "Group"))


# --- (STEP 3) Create the chart ---

# Define the display order for skills based on the Mann Whitney U table
# UPDATED: This order now matches your provided Mann-Whitney U table.
custom_skill_order_s8 <- c(
  "Feline closed extraction",
  "Canine open extraction - Single root",
  "Feline open extraction - Single root",
  "Canine open extraction - Canines/multiple roots",
  "Radiographic interpretation",
  "Feline open extraction - Canines/multiple roots", # Included again as it appears in your table
  "Canine closed extraction",
  "Radiographic positioning",
  "Scaling",
  "Polishing",
  "Oral exam/charting"
)

# Apply the custom order to the plot data.
# We reverse the order here because ggplot plots factors from bottom to top,
# so to have your custom order from top to bottom on the chart, you reverse it for the levels.
plot_data_s8$Skill <- factor(plot_data_s8$Skill, levels = rev(custom_skill_order_s8))

# Define the 4-color palette
likert_colors_s8 <- c(
  "1 - Strongly Disagree" = "#de2d26",
  "2 - Disagree" = "#fc9272",
  "3 - Agree" = "#a1d99b",
  "4 - Strongly Agree" = "#2ca25f"
)

# Generate the plot
ggplot(plot_data_s8, aes(y = Skill)) +
  geom_bar(data = . %>% filter(Group == "Educator"),
           aes(x = Proportion_Diverge, fill = Response_Factor), stat = "identity", width = 0.8) +
  geom_bar(data = . %>% filter(Group == "Employer"),
           aes(x = Proportion_Diverge, fill = Response_Factor), stat = "identity", width = 0.8) +
  geom_vline(xintercept = 0, color = "grey40", linetype = "solid") +
  geom_text(data = total_n_s8,
            aes(x = ifelse(Group == "Educator", -0.05, 0.05),
                label = paste0("N=", N),
                hjust = ifelse(Group == "Educator", 1, 0)),
            size = 3.5, color = "black") +
  scale_x_continuous(
    breaks = seq(-1, 1, 0.2),
    labels = function(x) paste0(abs(x * 100), "%"),
    limits = c(-1.1, 1.1)
  ) +
  scale_fill_manual(
    values = likert_colors_s8,
    drop = FALSE,
    name = "Competence Level",
    breaks = c("1 - Strongly Disagree", "2 - Disagree", "3 - Agree", "4 - Strongly Agree"),
    guide = guide_legend(reverse = FALSE)
  ) +
  labs(
    title = "Perceived Competence of Clinical Procedures",
    subtitle = "Distribution of Likert Scale Responses by Group (Educator vs. Employer)",
    x = "Proportion of Responses",
    y = "Clinical Procedure"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.grid.major.y = element_blank(),
    legend.title = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold", margin = margin(r = 10))
  ) +
  annotate("text", x = -0.7, y = Inf, label = "Educator", vjust = 1.5, size = 5, fontface = "bold") +
  annotate("text", x = 0.7, y = Inf, label = "Employer", vjust = 1.5, size = 5, fontface = "bold")
```

While the effect sizes for several procedures suggest that educators may place slightly more value on certain clinical extraction skills and radiographic interpretation than employers, these differences were not statistically significant. The negligible effect sizes for basic procedures such as polishing and oral exams indicate a strong consensus between the two groups regarding these skills. Figure 5.4 shows that the distribution of responses is broadly similar, and although some differences exist, they do not reach statistical significance to suggest meaningful divergence in perceptions of clinical competencies. Notably, the procedure "Feline Open Extraction – Canines/Multiple Roots" exhibits the greatest variability; however, this may be influenced by the smaller number of responses for that item.

# Discussion/Conclusion

## Interpretation of Results

The results across suggest that educators and employers are largely aligned in their perceptions of dental training and graduate competence, with no statistically significant differences observed in most areas. Both groups reported similar beliefs about which skills are taught preclinically and how many procedures should be completed during clinical training. However, a few notable trends emerged. Employers tended to place slightly more importance on teaching advanced extractions and expected more hands on experience during dentistry rotations, while educators were slightly more confident in graduates' ability to interpret radio-graphs. Although these differences were not always statistically significant, they point to subtle mismatches in expectations that could be addressed through improved communication and curriculum transparency.

The findings suggest a disconnect between educators and employers in how they perceive both student preparedness and the methods used to develop clinical skills. Educators consistently rated students as more proficient across a range of skills, indicating greater confidence in the effectiveness of current training approaches. On the other hand, employers were more cautious in their assessments, which could reflect different expectations, experiences with new graduates, or varying standards in clinical practice. The significant difference in views on instructional formats also highlights a potential misalignment in how each group understands or values different types of training. Additionally, the fact that some employers believed none of the instructional methods were being used raises concerns about transparency or awareness of educational practices. Overall, these differences point to a need for stronger collaboration between academic programs and the veterinary industry to ensure students are being prepared in ways that are both educationally sound and professionally relevant.

## Implications of the Study

The findings of this study have several important implications for veterinary education and the transition of new graduates into clinical practice. Overall, the results suggest that educators and small animal primary care employers are largely aligned in their expectations and beliefs about dental competencies but key perception gaps remain that may influence workforce readiness.

First, while both groups generally agreed on the importance of dental training and reported similar expectations about pre-clinical and clinical instruction, employers consistently underestimated the extent of instruction that educators reported providing. This misalignment may contribute to a lack of confidence in new graduates and increase reliance on on the job training. Improved transparency and communication between DVM programs and employers could help align expectations and strengthen trust in the educational process.

Second, differences in perceived importance of specific skills especially in areas like complex extractions and fluoride treatments highlight opportunities to review and potentially revise preclinical curriculum content. Employers placed greater emphasis on advanced skills than educators, suggesting a need to evaluate whether current curricula fully reflect the demands of entry level practice.

Third, the trend for employers to expect a higher number of completed procedures during clinical rotations, particularly in dentistry specific settings, points to a possible mismatch in expectations about hands on experience. Veterinary programs may consider reevaluating case requirements and ensuring that students receive sufficient procedural volume to meet clinical demands upon graduation.

Finally, these findings underscore the importance of ongoing collaboration between academia and clinical practice. By fostering dialogue between educators and employers, DVM programs can ensure that their graduates are not only technically competent, but also perceived as well prepared by those who hire and supervise them. Closing perception gaps may help reduce onboarding burdens, promote early career confidence, and ultimately enhance the quality of care delivered in small animal dentistry.


## Limitations

This study had several limitations that should be considered when interpreting the results. The sample sizes were relatively small, particularly for the employer group, which had only 13 respondents after data cleaning. This limited the statistical power of the analyses and may reduce the generalizability of the findings. In addition, the uneven group sizes restricted the types of statistical tests that could be used.

The recruitment process may have introduced sampling bias. Participants were recruited through professional networks and personal outreach, which can result in self selection bias. The employer sample was disproportionately composed of practice owners compared to national demographics, which could have skewed the results toward the perspectives of individuals with more authority or stronger opinions about training and graduate readiness.

There were also geographic and institutional limitations. Most employer responses came from practices located on the U.S. East Coast, while educator responses were more geographically diverse. This imbalance may limit the applicability of the findings to rural clinics or educational contexts outside of North America.

All data were self reported which introduces additional concerns such as variation in how participants interpret survey items. For example, the concept of “competency” may be understood differently across respondents, even if they selected similar response options on the Likert scale.

Lastly, missing data was another concern. Several participants did not complete the entire survey, and certain questions had high rates of missing or invalid responses. These gaps led to the exclusion of some data points and may have affected the robustness of the results for specific comparisons.


## Recommendations

To improve alignment between veterinary educators and employers, several strategies should be considered. Communication between veterinary schools and employers should be strengthened, as employers often underestimated the level of preclinical and clinical dental training reported by educators. Programs can address this perception gap by sharing curricular details and competency expectations through mechanisms such as externship evaluations, employer advisory boards, or structured graduate transition reports. Clinical exposure to dental procedures should also be enhanced, particularly in dentistry rotations, where employers tended to expect a higher number of completed procedures. Increasing hands on opportunities in core dental skills, especially extractions and radiograph interpretation, can better align training with workplace expectations.

Clarifying learning objectives and defining shared competency standards is another important step. By standardizing what constitutes competence in key skills, educators and employers can more consistently assess graduate preparedness. This is especially relevant for procedures like open extractions and radiographic interpretation. Additionally, perceived discrepancies in the importance of certain skills, such as fluoride treatment and multirooted feline extractions, should be discussed within curriculum committees and employer advisory groups. These conversations can help determine whether changes in instruction or clearer expectation setting are needed. Ongoing dialogue between veterinary programs and employers through stakeholder surveys, curriculum reviews, and alumni feedback is essential to keeping curricula responsive to the evolving demands of clinical practice.

## Summary of Key Findings

There was broad agreement between educators and employers regarding new graduate competence, curriculum content, and priorities for dental skills training. However, employers tended to slightly underestimate the extent of preclinical training that educators reported providing. Notably, employers placed significantly greater importance on teaching certain skills preclinically, specifically feline open extractions involving multiroot teeth and fluoride treatment. Additionally, employers expected a higher number of completed procedures during dentistry rotations, indicating stronger expectations for clinical repetition and hands on experience. Overall, most other comparisons between the groups revealed no statistically significant differences, with effect sizes generally small or negligible.

## Final Thoughts

This analysis highlights generally strong alignment between educators and employers on core dental competencies, but also reveals subtle perception gaps, particularly in the expected number of clinical procedures and the importance of teaching advanced extractions and fluoride treatment. Employers tended to underestimate the scope of preclinical training and expected more procedural experience during clinical rotations. These differences, though not always statistically significant, point to a need for improved communication and transparency between veterinary programs and the profession to ensure shared expectations and better prepare graduates for clinical practice.

# Appendix

